---
title: "Ext_Model_Calibration_BWHS"
output: html_document
date: "2025-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label="allow")
```

```{r chunk1, echo=FALSE, include=FALSE}
#Here you will load all required packages and functions

#functions
source("FILEPATH/Functions_ExistingModelValidation.R", local=knitr::knit_global())

#incidence rates and mortality rates for the calculator
source("FILEPATH/BWHS_incmortrates.R", local=knitr::knit_global())

if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("dplyr"))
}
if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("tidyverse"))
}
if(!require(lubridate)){
  install.packages("lubridate")
  suppress.warnings(library("lubridate"))
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  suppress.warnings(library("ggplot2"))
}
if(!require(gtools)){
  install.packages("gtools")
  suppress.warnings(library("gtools"))
}
```

```{r include=FALSE}
#Please make sure you are working from valdata and valdata2 datasets that were created from Step2.Create2datasets_BWHS.Rmd

#If number of relatives is NA set to 0
valdata1<-valdata %>%
  mutate(brcancersis=ifelse(is.na(brcancersis),0, brcancersis),
         brcancerdau=ifelse (is.na(brcancerdau), 0, brcancerdau),
         brcancerdad=ifelse(is.na(brcancerdad),0, brcancerdad),
         brcancermom=ifelse(is.na(brcancerdad),0, brcancerdad),
         fhisfstbcnr=ifelse(is.na(fhisfstbcnr), 0, fhisfstbcnr),
         fam1grbc50=ifelse(is.na(fam1grbc50),0, fam1grbc50),
         fhx_fdr_brca=ifelse(is.na(fhx_fdr_brca),0, fhx_fdr_brca))

#Need to create max # relatives if some, but not all, info is available
valdata1<-valdata1 %>%
  rowwise %>%
  mutate(familyhistorynum=as.numeric(sum(brcancersis,brcancerdau,brcancermom,brcancerdad)),
        familyhistorynum2=as.numeric(sum(fhisfstbcnr,brcancerdad)),
        familyhistorynum3=as.numeric(sum(fam1grbc50)),
        familyhistorynum4=as.numeric(sum(as.numeric(fhx_fdr_brca))),
        maxfamhx=pmax(familyhistorynum,familyhistorynum2, familyhistorynum3, familyhistorynum4))

```
### Check max famhx variable
```{r}
table(valdata1$maxfamhx)
```

```{r include=FALSE}
#This section sets up the data in the format for BWHS risk calculator

valdata1 <- valdata1 %>%
  mutate (BRCAFAM_NUM=ifelse(fhx_fdr_brca==0, 0,
                               ifelse(maxfamhx==1, 1,
                                ifelse(maxfamhx>1, 2, 3))),
            BRCAFAMLT50=ifelse(is.na(fam1grbc50), 2, fam1grbc50),
            
            PROSTFAM=2, #all unknown
            BILAT= 2, #"BILAT", "Both ovaries removed (bilateral oophorectomy)", choices = list("No" = 1, "Yes" = 0, "Unknown" = 2), selected = character(0)),
            age_end5y=age+5,
            AGEGE50 =ifelse(age >= 50, 1, 0))
```

```{r include=FALSE}
#This section continues set up the data in the format for BWHS risk calculator

valdata1 <- valdata1 %>%
  mutate(   HT_TOT_IN=height_in,
            
            LACT_NEVER_IND=ifelse(parous==1 & !is.na(parous) & !is.na(breastfeed) & breastfeed==1 , 0, 1),
            
            #create indicator variable for premenopausal status
            # groups unknown with postmenopausal
            MENOSTAT_F_PRE_IND= ifelse(as.numeric(meno_status2)==1 & !is.na(meno_status2), 1,0),
            
            # create indicator variable for BMI >= 30
            BMI_D_GE30_IND =ifelse(bmi >= 30 & !is.na(bmi), 1, 0),
            
            # create indicator variable for BMI at age 18 <= 25
            BMI18_LT25_IND = ifelse(bmi_earlyadult <= 25 & !is.na(bmi_earlyadult), 1, 0),
            
            #Assume all bilat oophorectomy are "no" bc we don't have this data- in calculator, unknown are assigned "no"
            BILAT_NO_IND=1, 
            # create indicator variable for breast biopsy
            # groups unknown with no
            BIOPVAR_EVER_IND=ifelse(biopsy_num >= 1 & !is.na(biopsy_num), 1, 0),
            
            # create indicator variable for 1 relative diagnosed with breast cancer after age 50
            BRCAFAM50_1_IND=ifelse(as.numeric(BRCAFAM_NUM) == 1 & as.numeric(BRCAFAMLT50) != 1 & !is.na(BRCAFAM_NUM) & !is.na(BRCAFAMLT50), 1, 0),	
            # create indicator variable for 2 relatives diagnosed with breast cancer or any before age 50
            BRCAFAM50_2_IND=ifelse(as.numeric(BRCAFAM_NUM)==2 & !is.na(BRCAFAM_NUM) | as.numeric(BRCAFAMLT50)==1 & !is.na(BRCAFAMLT50), 1, 0),
        
            # create indicator variable for age at menarche < 14
            # groups unknown with < 14
            MENARC_LT14_IND=ifelse(as.numeric(agemenarche) <14 |is.na(agemenarche), 1, 0),
            # create age/menarche interaction variable
            AGEGE50_MENARCGE14=age_ge50*(1 - MENARC_LT14_IND),
            # create indicator variable for family history of prostate cancer
            # groups unknown with no
            PROSTFAM_D_YES_IND=ifelse(as.numeric(PROSTFAM) == 1, 1, 0),
            # create age/biopsy interaction variable
            AGEGE50_BIOPVAR_EVER=age_ge50*BIOPVAR_EVER_IND,
            # create age/family history interaction variables
            AGEGE50_BRCAFAM50_1 =age_ge50 * BRCAFAM50_1_IND,
            AGEGE50_BRCAFAM50_2 =age_ge50 * BRCAFAM50_2_IND,
            # create BMI/menopause interaction variable
            BMIGE30_MENO_PRE = BMI_D_GE30_IND * MENOSTAT_F_PRE_IND,
            #1 is nulliparous here- assuming missing are nulliparous
            #changed 1.24.25
            NULPAR_PAROUS_IND=ifelse(parous==0 | is.na(parous), 0, 1),
            #1 is duration 5+y, 0 is none or <5y or unknown
            OCDUR_YRS_GE5_IND=ifelse(ocuse_dur/12>=5 & !is.na(ocuse_dur), 1, 0),
            
            AGEGE50_BRCAFAM50_1 = age_ge50* BRCAFAM50_1_IND,
            AGEGE50_BRCAFAM50_2 = age_ge50* BRCAFAM50_2_IND)

#Assign risk 

# 1 - the attributable risk 
BRCAN_1_AR <- 0.3576669608

# parameter estimates to calculate linear predictors
AGEGE50_BIOPVAR_EVER_BETA <- -0.241570698
AGEGE50_BRCAFAM50_1_BETA <- 0.0813301514
AGEGE50_BRCAFAM50_2_BETA <- -0.556421235
AGEGE50_MENARCGE14_BETA <- 0.2388811256
BILAT_NO_BETA <- 0.214297404
BIOPVAR_EVER_BETA <- 0.4629495248
BMI18_LT25_BETA <- 0.1508789583
BMIGE30_MENO_PRE_BETA <- -0.165478793
BMI_D_GE30_BETA <- 0.1284392293
BRCAFAM50_1_BETA <- 0.3416303435
BRCAFAM50_2_BETA <- 0.959041669
LACT_NEVER_BETA <- 0.1682571258
MENARC_LT14_BETA <- 0.3177481922
MENOSTAT_F_PRE_BETA <- 0.2383607104
NULPAR_PAROUS_BETA <- 0.0169183249
OCDUR_YRS_GE5_BETA <- 0.1398767985
PROSTFAM_D_YES_BETA <- 0.1363762353

# create linear predictor for each variable
valdata1<- valdata1 %>%
  mutate(LP_BILAT = BILAT_NO_IND * BILAT_NO_BETA,
         LP_BIOPVAR =BIOPVAR_EVER_IND*BIOPVAR_EVER_BETA + AGEGE50_BIOPVAR_EVER*AGEGE50_BIOPVAR_EVER_BETA,
         LP_BMI18 = BMI18_LT25_IND* BMI18_LT25_BETA,
         LP_BMI_BY_MENO = BMI_D_GE30_IND*BMI_D_GE30_BETA + MENOSTAT_F_PRE_IND*MENOSTAT_F_PRE_BETA + BMIGE30_MENO_PRE* BMIGE30_MENO_PRE_BETA,
         LP_BRCAFAM50 = BRCAFAM50_1_IND*BRCAFAM50_1_BETA + AGEGE50_BRCAFAM50_1*AGEGE50_BRCAFAM50_1_BETA + BRCAFAM50_2_IND * BRCAFAM50_2_BETA + AGEGE50_BRCAFAM50_2 * AGEGE50_BRCAFAM50_2_BETA,
         LP_LACTEVER_BIN = (as.numeric(LACT_NEVER_IND) * LACT_NEVER_BETA),
         LP_MENARC = MENARC_LT14_IND* MENARC_LT14_BETA + AGEGE50_MENARCGE14 * AGEGE50_MENARCGE14_BETA,
         LP_NULPAR = as.numeric(NULPAR_PAROUS_IND) * NULPAR_PAROUS_BETA,
         LP_OCDUR_YRS = as.numeric(OCDUR_YRS_GE5_IND) * OCDUR_YRS_GE5_BETA,
         LP_PROSTFAM_D = PROSTFAM_D_YES_IND* PROSTFAM_D_YES_BETA,
         LP_bwhs =LP_BILAT + LP_BIOPVAR + LP_BMI18 + LP_BMI_BY_MENO + LP_BRCAFAM50 + LP_LACTEVER_BIN + LP_MENARC + LP_NULPAR + LP_OCDUR_YRS + LP_PROSTFAM_D)

valdata1<-valdata1 %>%
    mutate(RRStar = exp(LP_bwhs))
```

### Check that no indiv. is missing RRStar

```{r }
#check that =0. If >0 there has been an error in the coding where the lp has not been set. Please check the names of all your variables to make sure they align
sum(is.na(valdata1$RRStar))
```

```{r include=FALSE}
#This section assigns risks for 1-10 y

#risks are stored as AbsRsk5 : 5y risk and AbsRsk10: 10y risk

valdata1 <- valdata1 %>%
  mutate(age=as.integer(age),
         t1=age,
         t2=t1+1,
         t3=t1+2,
         t4=t1+3,
         t5=t1+4,
         t6=t1+5,
         t7=t1+6,
         t8=t1+7,
         t9=t1+8,
         t10=t1+9, 
         
         H1 = BC_AA_inc[BC_AA_inc$age==t1,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t1,]$h2,
         H2 = BC_AA_inc[BC_AA_inc$age==t2,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t2,]$h2,
         H3 = BC_AA_inc[BC_AA_inc$age==t3,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t3,]$h2,
         H4 = BC_AA_inc[BC_AA_inc$age==t4,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t4,]$h2,
         H5 = BC_AA_inc[BC_AA_inc$age==t5,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t5,]$h2,
         
         H6 = BC_AA_inc[BC_AA_inc$age==t6,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t6,]$h2,
         H7 = BC_AA_inc[BC_AA_inc$age==t7,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t7,]$h2,
         H8 = BC_AA_inc[BC_AA_inc$age==t8,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t8,]$h2,
         H9 = BC_AA_inc[BC_AA_inc$age==t9,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t9,]$h2,
         H10 = BC_AA_inc[BC_AA_inc$age==t10,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t10,]$h2,
         
         Cum_H1 = 0,
         Cum_H2 = H1,
         Cum_H3 = H1 + H2,
         Cum_H4 = H1 + H2 + H3,
         Cum_H5 = H1 + H2 + H3 + H4,
         
         Cum_H6 = H1 + H2 + H3 + H4 + H5,
         Cum_H7 = H1 + H2 + H3 + H4 + H5 + H6,
         Cum_H8 = H1 + H2 + H3 + H4 + H5 + H6 + H7,
         Cum_H9 = H1 + H2 + H3 + H4 + H5 + H6 + H7 + H8,
         Cum_H10 = H1 + H2 + H3 + H4 + H5 + H6 + H7 + H8 + H9,
         
         PI_1 = (( (BC_AA_inc[BC_AA_inc$age==t1,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H1 ) * exp(-Cum_H1)) * (1-exp(-H1)),
         PI_2 = (( (BC_AA_inc[BC_AA_inc$age==t2,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H2 ) * exp(-Cum_H2)) * (1-exp(-H2)),
         PI_3 = (( (BC_AA_inc[BC_AA_inc$age==t3,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H3 ) * exp(-Cum_H3)) * (1-exp(-H3)),
         PI_4 = (( (BC_AA_inc[BC_AA_inc$age==t4,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H4 ) * exp(-Cum_H4)) * (1-exp(-H4)),
         PI_5 = (( (BC_AA_inc[BC_AA_inc$age==t5,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H5 ) * exp(-Cum_H5)) * (1-exp(-H5)),
         
         PI_6 = (( (BC_AA_inc[BC_AA_inc$age==t6,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H6 ) * exp(-Cum_H6)) * (1-exp(-H6)),
         PI_7 = (( (BC_AA_inc[BC_AA_inc$age==t7,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H7 ) * exp(-Cum_H7)) * (1-exp(-H7)),
         PI_8 = (( (BC_AA_inc[BC_AA_inc$age==t8,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H8 ) * exp(-Cum_H8)) * (1-exp(-H8)),
         PI_9 = (( (BC_AA_inc[BC_AA_inc$age==t9,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H9 ) * exp(-Cum_H9)) * (1-exp(-H9)),
         PI_10 = (( (BC_AA_inc[BC_AA_inc$age==t10,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H10 ) * exp(-Cum_H10)) * (1-exp(-H10)),
         
         
         AbsRsk5 = round(100*(PI_1 + PI_2 + PI_3 + PI_4 + PI_5), 1),
         AbsRsk10 = round(100*(PI_1 + PI_2 + PI_3 + PI_4 + PI_5 + PI_6 + PI_7 + PI_8 + PI_9 + PI_10), 1),
         AvAbsRsk5 = list(round(average[average$age==t1,]$predrisk5, 1)),
         AvAbsRsk10 = list(round(average[average$age==t1,]$predrisk10, 1)),
         absrisk=AbsRsk5/100
  )

valdata1<-as.data.frame(valdata1)
```

## Risk summary, BWHS 1-5y

```{r echo=FALSE}
summarizerisks<-function(data, var){
  data %>% group_by(cohort) %>%
    summarise(n= sum(!is.na({{var}})),
              na= sum(is.na({{var}})),
              mean = mean({{var}}, na.rm = TRUE),
              sd = sd({{var}}, na.rm = TRUE),
              min=min({{var}}, na.rm=TRUE),
              q25 = quantile({{var}}, 0.25, na.rm = TRUE),
              median=median({{var}}, na.rm=TRUE),
              q75 = quantile({{var}}, 0.75, na.rm = TRUE),
              max= max({{var}}, na.rm=TRUE))
}

a<-summarizerisks(valdata1, RRStar)
b<-summarizerisks(valdata1, absrisk)
bwhsrisks<-rbind(a,b)
bwhsrisks
hist(valdata1$RRStar, col="darkblue")
hist(valdata1$absrisk, col="darkred")
```
```{r}
#OPTIONAL CODE: 
#valdata1_risks<-valdata1[,c("cohort","subject_id","RRStar","absrisk","AbsRsk5","AbsRsk10")]
#Save absolute risks 
#write.csv(valdata1_risks, "C:/Users/krist/Box/BCRPP Concept 2_ValidatingExistingModels/Data - Opted-in Studies/AllRisks/1to5y/BWHS/WHI.csv")
```

```{r include=FALSE}


########################################################
# Calibration - use iCARE Validate    #
########################################################

#timevar needs to be changed to meet requirements, change to infinity if not a case (invasive or in situ).

#If timediff case yr is missing assign as infinite (past 5 y)
#Reassigned so all that are =5y are considered "within" 5 y (conversation with Julie and Gary 1.24.25)
valdata1$time.of.onset<-ifelse(valdata1$case_invasive==1 & !is.na(valdata1$time_bltodx1) & valdata1$time_bltodx1<=5, valdata1$time_bltodx1, Inf)
valdata1$case_5yinv<-ifelse(valdata1$case_invasive==1 & valdata1$time_bltodx1<=5, 1, 0)

table(valdata1$case_5yinv, valdata1$time_bltodx1)

#Add age end
valdata1<-valdata1%>%
  mutate(sampling.fraction=1,
         ageend=pmin(age+5, age+timediff))

#5y invasive first 
#Rename

valdata1_new<-valdata1[,c("id","age","ageend","case_5yinv","RRStar","absrisk","time.of.onset","sampling.fraction", "LP_bwhs","race", "cohort")] 

#Rename columns so it matches iCARE scheme
colnames(valdata1_new)<-c("id","study.entry.age","study.exit.age","observed.outcome","relrisk","absrisk", "time.of.onset","sampling.weights","lp", "race","cohort")
```

## E/O overall 1-5y

```{r echo=FALSE}

eo<-as.data.frame(obsexpected_overall(valdata1_new))
eo
```
## E/O by decile 1-5y

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(valdata1_new))
print("E/O by cat, invasive")
eo_bycat_invas
```

## Calibration plots 1-5y

```{r echo=FALSE}
absrisk_byabs<-as.data.frame(returnabsrisk_abs(valdata1_new))
print("Absolute risk by absolute risk decile")
absrisk_byabs

plotcalibration_abs= function(dataset1){
  maxx=max(dataset1$predicted.prob.cat)*100+0.2
  maxy=max(dataset1$upper.limit.absrisk.cat)*100+0.2
  miny=min(dataset1$lower.limit.absrisk.cat)*100-0.2
  plot_test<-ggplot(data=dataset1,aes(x=predicted.prob.cat*100,y=observed.prob.cat*100)) +
    geom_point(color="purple4",size=4) +
    geom_abline(intercept=0, slope=1,linetype="dotted") + geom_errorbar(aes(ymin=lower.limit.absrisk.cat*100, ymax=upper.limit.absrisk.cat*100), size=0.3) +
    labs(x='Expected risk (%)', y='Observed risk (%)', title="Absolute risk calibration")+ 
    scale_x_continuous(limits=c(0,maxx))  + scale_y_continuous(limits=c(miny,maxy)) +
    theme(axis.text.x=element_text(size=12),
          axis.text.y=element_text(size=12),
          axis.title.x = element_text(vjust = 0, size = 14),
          axis.title.y = element_text(vjust = 2, size = 14), 
          axis.title=element_text(size=16), legend.position="none") 
  return(plot_test)
}      
a2<-suppressWarnings(plotcalibration_abs(absrisk_byabs))
a2

```

## AUC results 1-5y

## AUC relative risk

```{r}
AUC_relrisk<-as.data.frame(AUCcalc(valdata1_new))
AUC_relrisk
```

## AUC absolute risk 1-5y
```{r}

AUC_absrisk<-as.data.frame(AUCcalc_absrisk(valdata1_new))
print("AUC by absolute risk")
AUC_absrisk
```

## AUC absolute risk age adjusted 1-5y
```{r}

AUC_absrisk_ageadj<-as.data.frame(AUCcalc_absrisk_ageadjust(valdata1_new))
AUC_absrisk_ageadj
```

```{r include=FALSE}

#Datasetup for 2-6y
#########################
# For 2-6y
#first make sure nobody >70 is included

valdata2<-valdata2[which (valdata2$age<=70),]

table(valdata$case_6y)
table(valdata2$case_6yinv)
rm(valdata)
#If number of relatives is NA set to 0
valdata<-valdata2 %>%
  mutate(brcancersis=ifelse(is.na(brcancersis),0, brcancersis),
         brcancerdau=ifelse (is.na(brcancerdau), 0, brcancerdau),
         brcancerdad=ifelse(is.na(brcancerdad),0, brcancerdad),
         brcancermom=ifelse(is.na(brcancerdad),0, brcancerdad),
         fhisfstbcnr=ifelse(is.na(fhisfstbcnr), 0, fhisfstbcnr),
         fam1grbc50=ifelse(is.na(fam1grbc50),0, fam1grbc50),
         fhx_fdr_brca=ifelse(is.na(fhx_fdr_brca),0, fhx_fdr_brca))

#Need to create max # relatives if some, but not all, info is available
valdata<-valdata %>%
  rowwise %>%
  mutate(familyhistorynum=as.numeric(sum(brcancersis,brcancerdau,brcancermom,brcancerdad)),
         familyhistorynum2=as.numeric(sum(fhisfstbcnr,brcancerdad)),
         familyhistorynum3=as.numeric(sum(fam1grbc50)),
         familyhistorynum4=as.numeric(sum(as.numeric(fhx_fdr_brca))),
         maxfamhx=pmax(familyhistorynum,familyhistorynum2, familyhistorynum3, familyhistorynum4))

valdata <- valdata %>%
  mutate (BRCAFAM_NUM=ifelse(fhx_fdr_brca==0, 0,
                             ifelse(maxfamhx==1, 1,
                                    ifelse(maxfamhx>1, 2, 3))),
          BRCAFAMLT50=ifelse(is.na(fam1grbc50), 2, fam1grbc50),
          
          PROSTFAM=2, #all unknown
          BILAT= 2, #"BILAT", "Both ovaries removed (bilateral oophorectomy)", choices = list("No" = 1, "Yes" = 0, "Unknown" = 2), selected = character(0)),
          age_end5y=age+5,
          AGEGE50 =ifelse(age >= 50, 1, 0))

valdata <- valdata %>%
  mutate(   HT_TOT_IN=height_in,
            
            LACT_NEVER_IND=ifelse(parous==1 & !is.na(parous) & !is.na(breastfeed) & breastfeed==1 , 0, 1),
            
            #create indicator variable for premenopausal status
            # groups unknown with postmenopausal
            MENOSTAT_F_PRE_IND= ifelse(as.numeric(meno_status2)==1 & !is.na(meno_status2), 1,0),
            
            # create indicator variable for BMI >= 30
            BMI_D_GE30_IND =ifelse(bmi >= 30 & !is.na(bmi), 1, 0),
            
            # create indicator variable for BMI at age 18 <= 25
            BMI18_LT25_IND = ifelse(bmi_earlyadult <= 25 & !is.na(bmi_earlyadult), 1, 0),
            
            #Assume all bilat oophorectomy are "no" bc we don't have this data- in calculator, unknown are assigned "no"
            BILAT_NO_IND=1, 
            # create indicator variable for breast biopsy
            # groups unknown with no
            BIOPVAR_EVER_IND=ifelse(biopsy_num >= 1 & !is.na(biopsy_num), 1, 0),
            
            # create indicator variable for 1 relative diagnosed with breast cancer after age 50
            BRCAFAM50_1_IND=ifelse(as.numeric(BRCAFAM_NUM) == 1 & as.numeric(BRCAFAMLT50) != 1 & !is.na(BRCAFAM_NUM) & !is.na(BRCAFAMLT50), 1, 0),	
            # create indicator variable for 2 relatives diagnosed with breast cancer or any before age 50
            BRCAFAM50_2_IND=ifelse(as.numeric(BRCAFAM_NUM)==2 & !is.na(BRCAFAM_NUM) | as.numeric(BRCAFAMLT50)==1 & !is.na(BRCAFAMLT50), 1, 0),
            
            # create indicator variable for age at menarche < 14
            # groups unknown with < 14
            MENARC_LT14_IND=ifelse(as.numeric(agemenarche) <14 |is.na(agemenarche), 1, 0),
            # create age/menarche interaction variable
            AGEGE50_MENARCGE14=age_ge50*(1 - MENARC_LT14_IND),
            # create indicator variable for family history of prostate cancer
            # groups unknown with no
            PROSTFAM_D_YES_IND=ifelse(as.numeric(PROSTFAM) == 1, 1, 0),
            # create age/biopsy interaction variable
            AGEGE50_BIOPVAR_EVER=age_ge50*BIOPVAR_EVER_IND,
            # create age/family history interaction variables
            AGEGE50_BRCAFAM50_1 =age_ge50 * BRCAFAM50_1_IND,
            AGEGE50_BRCAFAM50_2 =age_ge50 * BRCAFAM50_2_IND,
            # create BMI/menopause interaction variable
            BMIGE30_MENO_PRE = BMI_D_GE30_IND * MENOSTAT_F_PRE_IND,
            #1 is nulliparous here- assuming missing are nulliparous
            #changed 1.24.25
            NULPAR_PAROUS_IND=ifelse(parous==0 | is.na(parous), 0, 1),
            #1 is duration 5+y, 0 is none or <5y or unknown
            OCDUR_YRS_GE5_IND=ifelse(ocuse_dur/12>=5 & !is.na(ocuse_dur), 1, 0),
            
            AGEGE50_BRCAFAM50_1 = age_ge50* BRCAFAM50_1_IND,
            AGEGE50_BRCAFAM50_2 = age_ge50* BRCAFAM50_2_IND)

#Assign risk 

# 1 - the attributable risk 
BRCAN_1_AR <- 0.3576669608

# parameter estimates to calculate linear predictors
AGEGE50_BIOPVAR_EVER_BETA <- -0.241570698
AGEGE50_BRCAFAM50_1_BETA <- 0.0813301514
AGEGE50_BRCAFAM50_2_BETA <- -0.556421235
AGEGE50_MENARCGE14_BETA <- 0.2388811256
BILAT_NO_BETA <- 0.214297404
BIOPVAR_EVER_BETA <- 0.4629495248
BMI18_LT25_BETA <- 0.1508789583
BMIGE30_MENO_PRE_BETA <- -0.165478793
BMI_D_GE30_BETA <- 0.1284392293
BRCAFAM50_1_BETA <- 0.3416303435
BRCAFAM50_2_BETA <- 0.959041669
LACT_NEVER_BETA <- 0.1682571258
MENARC_LT14_BETA <- 0.3177481922
MENOSTAT_F_PRE_BETA <- 0.2383607104
NULPAR_PAROUS_BETA <- 0.0169183249
OCDUR_YRS_GE5_BETA <- 0.1398767985
PROSTFAM_D_YES_BETA <- 0.1363762353

# create linear predictor for each variable
valdata<- valdata %>%
  mutate(LP_BILAT = BILAT_NO_IND * BILAT_NO_BETA,
         LP_BIOPVAR =BIOPVAR_EVER_IND*BIOPVAR_EVER_BETA + AGEGE50_BIOPVAR_EVER*AGEGE50_BIOPVAR_EVER_BETA,
         LP_BMI18 = BMI18_LT25_IND* BMI18_LT25_BETA,
         LP_BMI_BY_MENO = BMI_D_GE30_IND*BMI_D_GE30_BETA + MENOSTAT_F_PRE_IND*MENOSTAT_F_PRE_BETA + BMIGE30_MENO_PRE* BMIGE30_MENO_PRE_BETA,
         LP_BRCAFAM50 = BRCAFAM50_1_IND*BRCAFAM50_1_BETA + AGEGE50_BRCAFAM50_1*AGEGE50_BRCAFAM50_1_BETA + BRCAFAM50_2_IND * BRCAFAM50_2_BETA + AGEGE50_BRCAFAM50_2 * AGEGE50_BRCAFAM50_2_BETA,
         LP_LACTEVER_BIN = (as.numeric(LACT_NEVER_IND) * LACT_NEVER_BETA),
         LP_MENARC = MENARC_LT14_IND* MENARC_LT14_BETA + AGEGE50_MENARCGE14 * AGEGE50_MENARCGE14_BETA,
         LP_NULPAR = as.numeric(NULPAR_PAROUS_IND) * NULPAR_PAROUS_BETA,
         LP_OCDUR_YRS = as.numeric(OCDUR_YRS_GE5_IND) * OCDUR_YRS_GE5_BETA,
         LP_PROSTFAM_D = PROSTFAM_D_YES_IND* PROSTFAM_D_YES_BETA,
         LP_bwhs =LP_BILAT + LP_BIOPVAR + LP_BMI18 + LP_BMI_BY_MENO + LP_BRCAFAM50 + LP_LACTEVER_BIN + LP_MENARC + LP_NULPAR + LP_OCDUR_YRS + LP_PROSTFAM_D)

valdata<-valdata %>%
  mutate(RRStar = exp(LP_bwhs))
```

### Check that no missing RRStar 2-6y
```{r}
#check that =0
sum(is.na(valdata$RRStar))
```

```{r include=FALSE}
#Assign risks
#risks are stored as AbsRsk5 : 5y risk and AbsRsk10: 10y risk

valdata <- valdata %>%
  mutate(age=as.integer(age),
         t1=age,
         t2=t1+1,
         t3=t1+2,
         t4=t1+3,
         t5=t1+4,
         t6=t1+5,
         t7=t1+6,
         t8=t1+7,
         t9=t1+8,
         t10=t1+9, 
         
         H1 = BC_AA_inc[BC_AA_inc$age==t1,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t1,]$h2,
         H2 = BC_AA_inc[BC_AA_inc$age==t2,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t2,]$h2,
         H3 = BC_AA_inc[BC_AA_inc$age==t3,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t3,]$h2,
         H4 = BC_AA_inc[BC_AA_inc$age==t4,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t4,]$h2,
         H5 = BC_AA_inc[BC_AA_inc$age==t5,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t5,]$h2,
         
         H6 = BC_AA_inc[BC_AA_inc$age==t6,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t6,]$h2,
         H7 = BC_AA_inc[BC_AA_inc$age==t7,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t7,]$h2,
         H8 = BC_AA_inc[BC_AA_inc$age==t8,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t8,]$h2,
         H9 = BC_AA_inc[BC_AA_inc$age==t9,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t9,]$h2,
         H10 = BC_AA_inc[BC_AA_inc$age==t10,]$BRCAN_inc * BRCAN_1_AR * RRStar + mort[mort$age==t10,]$h2,
         
         Cum_H1 = 0,
         Cum_H2 = H1,
         Cum_H3 = H1 + H2,
         Cum_H4 = H1 + H2 + H3,
         Cum_H5 = H1 + H2 + H3 + H4,
         
         Cum_H6 = H1 + H2 + H3 + H4 + H5,
         Cum_H7 = H1 + H2 + H3 + H4 + H5 + H6,
         Cum_H8 = H1 + H2 + H3 + H4 + H5 + H6 + H7,
         Cum_H9 = H1 + H2 + H3 + H4 + H5 + H6 + H7 + H8,
         Cum_H10 = H1 + H2 + H3 + H4 + H5 + H6 + H7 + H8 + H9,
         
         PI_1 = (( (BC_AA_inc[BC_AA_inc$age==t1,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H1 ) * exp(-Cum_H1)) * (1-exp(-H1)),
         PI_2 = (( (BC_AA_inc[BC_AA_inc$age==t2,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H2 ) * exp(-Cum_H2)) * (1-exp(-H2)),
         PI_3 = (( (BC_AA_inc[BC_AA_inc$age==t3,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H3 ) * exp(-Cum_H3)) * (1-exp(-H3)),
         PI_4 = (( (BC_AA_inc[BC_AA_inc$age==t4,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H4 ) * exp(-Cum_H4)) * (1-exp(-H4)),
         PI_5 = (( (BC_AA_inc[BC_AA_inc$age==t5,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H5 ) * exp(-Cum_H5)) * (1-exp(-H5)),
         
         PI_6 = (( (BC_AA_inc[BC_AA_inc$age==t6,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H6 ) * exp(-Cum_H6)) * (1-exp(-H6)),
         PI_7 = (( (BC_AA_inc[BC_AA_inc$age==t7,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H7 ) * exp(-Cum_H7)) * (1-exp(-H7)),
         PI_8 = (( (BC_AA_inc[BC_AA_inc$age==t8,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H8 ) * exp(-Cum_H8)) * (1-exp(-H8)),
         PI_9 = (( (BC_AA_inc[BC_AA_inc$age==t9,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H9 ) * exp(-Cum_H9)) * (1-exp(-H9)),
         PI_10 = (( (BC_AA_inc[BC_AA_inc$age==t10,]$BRCAN_inc * BRCAN_1_AR * RRStar) / H10 ) * exp(-Cum_H10)) * (1-exp(-H10)),
         
         
         AbsRsk5 = round(100*(PI_1 + PI_2 + PI_3 + PI_4 + PI_5), 1),
         AbsRsk10 = round(100*(PI_1 + PI_2 + PI_3 + PI_4 + PI_5 + PI_6 + PI_7 + PI_8 + PI_9 + PI_10), 1),
         AvAbsRsk5 = round(average[average$age==t1,]$predrisk5, 1),
         AvAbsRsk10 = round(average[average$age==t1,]$predrisk10, 1),
         absrisk=AbsRsk5/100
  )

valdata<-as.data.frame(valdata)

```

## Summary risks 2-6y

```{r echo=FALSE}
summarizerisks<-function(data, var){
  data %>% group_by(cohort) %>%
    summarise(n= sum(!is.na({{var}})),
              na= sum(is.na({{var}})),
              mean = mean({{var}}, na.rm = TRUE),
              sd = sd({{var}}, na.rm = TRUE),
              min=min({{var}}, na.rm=TRUE),
              q25 = quantile({{var}}, 0.25, na.rm = TRUE),
              median=median({{var}}, na.rm=TRUE),
              q75 = quantile({{var}}, 0.75, na.rm = TRUE),
              max= max({{var}}, na.rm=TRUE))
}

a<-summarizerisks(valdata, RRStar)
b<-summarizerisks(valdata, absrisk)
bwhsrisks<-rbind(a,b)
bwhsrisks
hist(valdata$RRStar, col="darkblue")
hist(valdata$absrisk, col="darkred")
valdata_risks<-valdata[,c("cohort","subject_id","RRStar","absrisk","AbsRsk5","AbsRsk10")]
```

```{r include=FALSE}
#OPTIONAL CODE:
#Save absolute risks 
#write.csv(valdata_risks, "C:/Users/krist/Box/BCRPP Concept 2_ValidatingExistingModels/Data - Opted-in Studies/AllRisks/2to6y/BWHS/WHI.csv")
```

## Calibration 2-6y
```{r include=FALSE}
########################################################
# Calibration - use iCARE Validate    #
########################################################

#timevar needs to be changed to meet requirements, change to infinity if not a case (invasive or in situ).

#If timediff case yr is missing assign as infinite (past 5 y)
#Reassigned so all that are =5y are considered "within" 5 y (conversation with Julie and Gary 1.24.25)
valdata$time.of.onset<-ifelse(valdata$case_invasive==1 & !is.na(valdata$time_bltodx1) & valdata$time_bltodx1<=5, valdata$time_bltodx1, Inf)
valdata$case_5yinv<-ifelse(valdata$case_invasive==1 & valdata$time_bltodx1<=5, 1, 0)


#Add age end
valdata<-valdata%>%
  mutate(sampling.fraction=1,
         ageend=pmin(age+5, age+timediff))

#5y invasive first 
#Rename

valdata_new<-valdata[,c("id","age","ageend","case_5yinv","RRStar","absrisk","time.of.onset","sampling.fraction", "LP_bwhs","race", "cohort")] 

#Rename columns so it matches iCARE scheme
colnames(valdata_new)<-c("id","study.entry.age","study.exit.age","observed.outcome","relrisk","absrisk", "time.of.onset","sampling.weights","lp", "race","cohort")
```

## E/O overall 2-6y
```{r}
eo<-as.data.frame(obsexpected_overall(valdata_new))
eo
```
## E/O by decile 2-6y
```{r}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(valdata_new))
print("E/O by cat, invasive")
eo_bycat_invas
```
## Calibration plots
```{r echo=FALSE}
absrisk_byabs<-as.data.frame(returnabsrisk_abs(valdata_new))
print("Absolute risk by absolute risk decile")
absrisk_byabs

plotcalibration_abs= function(dataset1){
  maxx=max(dataset1$predicted.prob.cat)*100+0.2
  maxy=max(dataset1$upper.limit.absrisk.cat)*100+0.2
  miny=min(dataset1$lower.limit.absrisk.cat)*100-0.2
  plot_test<-ggplot(data=dataset1,aes(x=predicted.prob.cat*100,y=observed.prob.cat*100)) +
    geom_point(color="purple4",size=4) +
    geom_abline(intercept=0, slope=1,linetype="dotted") + geom_errorbar(aes(ymin=lower.limit.absrisk.cat*100, ymax=upper.limit.absrisk.cat*100), size=0.3) +
    labs(x='Expected risk (%)', y='Observed risk (%)', title="Absolute risk calibration")+ 
    scale_x_continuous(limits=c(0,maxx))  + scale_y_continuous(limits=c(miny,maxy)) +
    theme(axis.text.x=element_text(size=12),
          axis.text.y=element_text(size=12),
          axis.title.x = element_text(vjust = 0, size = 14),
          axis.title.y = element_text(vjust = 2, size = 14), 
          axis.title=element_text(size=16), legend.position="none") 
  return(plot_test)
}      
a2<-suppressWarnings(plotcalibration_abs(absrisk_byabs))
a2

```
## AUC for 2-6y

## AUC rel risk

```{r}
# AUC 
AUC_relrisk<-as.data.frame(AUCcalc(valdata_new))
AUC_relrisk
```

## AUC by absolute risk

```{r}
AUC_absrisk<-as.data.frame(AUCcalc_absrisk(valdata_new))
print("AUC by absolute risk")
AUC_absrisk

```
## AUC by absolute risk age adjusted
```{r}
AUC_absrisk_ageadj<-as.data.frame(AUCcalc_absrisk_ageadjust(valdata_new))
AUC_absrisk_ageadj
```

```{r instructions, include=FALSE}
#You will now need to run rmarkdown::render in your R Console. This cannot be included in the .Rmd file directly, or it will iteratively run.
#Please enter this code into the console:
#     rmarkdown::render("FILEPATH/Ext_Model_Calibration_BWHS.Rmd")
#The resulting .html file will be called Ext_Model_Calibration_Gail as well, and will be found in the same file path as that of your .Rmd file.
```
