```{r include=FALSE}
#Instructions:

# This file will read in your clean validation dataset, created after
# running "Step1.DataCleaning.R" and will perform model calibration
# using iCARE-Validate for the Gail (BCRAT) model.

# PLEASE INSERT COHORT ACRONYM IN [] BELOW for the FILENAME
# AND INSERT RUN DATE



# Run Notes: You will need to install these packages if you don't have them (available online):
# #BCRAT packages - install.packages("BCRA") - please run outside of Rmarkdown:
#  https://dceg.cancer.gov/tools/risk-assessment/bcra
# #iCARE packages - install.packages("BiocManager")
  #                  BiocManager::install("iCARE")
# https://www.bioconductor.org/packages/release/bioc/html/iCARE.html

#Dataset: You will use datasets from Create2datasets.Rmd: valdata and valdata2. 
#If these are not in your R environment, please re-run Step2.Create2datasets.Rmd without knitting, and do not clear your R environment.

# After specifying your filepaths, (ctrl+f: "filepath" for where to insert), you can simply:
# hit "Knit -> Knit to Word" and this will run the full document and produce output.

# Please send the finalized document to kristen_brantley@dfci.harvard.edu
# If any issues arise when trying to run this document, please reach out to Kristen Brantley. 
```


---
title: "Ext_Model_Calibration_Gail [CTS]"
output: html_document
date: "MM-DD-YYYY"
---
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label="allow")
```




```{r chunk1, echo=FALSE, include=FALSE}
#Here you will load all required packages and functions
source("FILEPATH/Functions_ExistingModelValidation.R", local=knitr::knit_global())

if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("dplyr"))
}
if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("tidyverse"))
}
if(!require(lubridate)){
  install.packages("lubridate")
  suppress.warnings(library("lubridate"))
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  suppress.warnings(library("ggplot2"))
}
if(!require(gtools)){
  install.packages("gtools")
  suppress.warnings(library("gtools"))
}

if(!require(BCRA)){
  install.packages("BCRA")
  suppress.warnings(library("BCRA"))
}
#Load the data from BCRA
data("BrCa_lambda1")
data("BrCa_lambda2")
data("BrCa_beta")
```

```{r instructions, include=FALSE}
#Note: If you need to explore the data set up further, 
 #         this will read in the example data from BCRAT tool (remove #)
          
          #```{r}
          
          #data(exampledata, package="BCRA")
          #mean(exampledata$RR_Star1)
          #check.summary(exampledata)
          #```
```          

```{r instructions, include=FALSE}
#Create variables for use in Gail Model 
#Note: In Gail model, all missing variables are set to the most common
```

```{r chunk4, echo = FALSE}
#Here, we restrict our datasets (valdata and valdata2) to include only variables needed for the Gail model and for calibration.

data_gail1 <-valdata[,c("subject_id","id","age","race","ethnicity","agemenarche","ageflb","famhx_first","biopsy_num","atyp_hyp","parous","cohort", "case_5yinv", "case_5y","timediff")]

data_gail2 <-valdata2[,c("subject_id","id","age","race","ethnicity","agemenarche","ageflb","famhx_first","biopsy_num","atyp_hyp","parous","cohort", "case_6yinv", "case_6y","timediff")]


```

```{r instructions, include=FALSE}   
#Reassign values based on Gail parameters
        # Missing data assignments
        # The standards used in the online Gail model were used for variables with missing data. 
        # Women with unknown age at menarche were assigned menarche at 14 years of age or older. 
        # Women with unknown age at first live birth were classified as giving birth before age 20 years. 
        # Women with missing family history were classified as having no family history. 
        # To verify agreement between the code we used and the online tool, we randomly selected 10 patients 
        # and compared the 5-year and lifetime risk estimates obtained from the code given us to those from the
        # online risk assessment tool. 
        # 
        # For our purposes, assign all Asian under "Chinese" for abs.risk calibration
        # Consider all hispanics american born
        # 1=Wh White 1983-87 SEER rates (rates used in NCI BCRAT)
        # 2=AA African-American
        # 3=HU Hispanic-American (US born) 1995-04
        # 4=NA Other (Native American and unknown race)
        # 5=HF Hispanic-American (Foreign born) 1995-04
        # 6=Ch Chinese-American
        # 7=Ja Japanese-American
        # 8=Fi Filipino-American
        # 9=Hw Hawaiian-American
        # 10=oP Other Pacific Islander
        # 11=oA Other Asian

       # Set end age as age+5 and +10 years
```

```{r chunk6, include=FALSE}
data_gail1 <- data_gail1 %>%
  mutate(gailrace=ifelse(race==1 & ethnicity==1, 3, ifelse(race==1 & ethnicity !=1 ,1, ifelse(race==2, 2, ifelse(race==5, 4, ifelse(race==3, 6,    ifelse(race==4, 9, 4)))))),
         gailrace=replace(gailrace, is.na(gailrace), 4),
         gailagemen=ifelse(is.na(agemenarche), 99, agemenarche),
         gailageflb=replace(ageflb, is.na(ageflb) & parous==0, 98),
         gailageflb=replace(gailageflb, is.na(gailageflb), 99),
         gailageflb=replace(gailageflb, gailageflb<gailagemen, 99), #set to missing if less than age at menarche
         gailnrels=ifelse(is.na(famhx_first), 99, famhx_first),
         gailbiop=ifelse(is.na(biopsy_num), 99, biopsy_num),
         gailatyp=ifelse(is.na(atyp_hyp), 99, atyp_hyp),
         #Brca projection age 
         gailageend5y=age+5,
         sampling.fraction=1)


data_gail2<-data_gail2 %>%
  mutate(gailrace=ifelse(race==1 & ethnicity==1, 3, ifelse(race==1 & ethnicity !=1 ,1, ifelse(race==2, 2,   ifelse(race==5, 4, ifelse(race==3, 6,    ifelse(race==4, 9, 4)))))),
         gailrace=replace(gailrace, is.na(gailrace), 4),
         gailagemen=ifelse(is.na(agemenarche), 99, agemenarche),
         gailageflb=replace(ageflb, is.na(ageflb) & parous==0, 98),
         gailageflb=replace(gailageflb, is.na(gailageflb), 99),
         gailageflb=replace(gailageflb, gailageflb<gailagemen, 99), 
         gailnrels=ifelse(is.na(famhx_first), 99, famhx_first),
         gailbiop=ifelse(is.na(biopsy_num), 99, biopsy_num),
         gailatyp=ifelse(is.na(atyp_hyp), 99, atyp_hyp),
         gailageend5y=age+5,
         sampling.fraction=1)
```

```{r chunk7, echo = T, results = 'hide'}
# Here we keep only variables needed for Gail risk prediction and calibration, from the new vars created.

#Years 1-5
data_gail1<-data_gail1[,c("subject_id","cohort","age","gailageend5y","gailrace","gailagemen","gailageflb","gailnrels","gailbiop","gailatyp", "case_5yinv", "case_5y", "timediff", "sampling.fraction")]
#Change names
names(data_gail1)<-c("id","cohort","T1","T2","Race","AgeMen","Age1st","N_Rels","N_Biop","HypPlas","caseinv", "case_any", "timevar","sampling.weights")

#Years 2-6
data_gail2<-data_gail2[,c("subject_id","cohort","age","gailageend5y","gailrace","gailagemen","gailageflb","gailnrels","gailbiop","gailatyp", "case_6yinv", "case_6y", "timediff","sampling.fraction")]
#Change names
names(data_gail2)<-c("id","cohort","T1","T2","Race","AgeMen","Age1st","N_Rels","N_Biop","HypPlas","caseinv", "case_any", "timevar","sampling.weights")

```

```{r chunk8 }
#Check your dataset!

recodechecktable<-recode.check(data_gail1, Raw_Ind=1)
test<-recodechecktable[which (recodechecktable$Error_Ind>0),]
test
      
#If any observations show up on this print out then these have errors 
#that need to be fixed in the assignment of raw gail vars.
      
#repeat check for gail data2
recodechecktable<-recode.check(data_gail2, Raw_Ind=1)
test2<-recodechecktable[which (recodechecktable$Error_Ind>0),]
test2
      
```


## Estimate Relative risk and absolute risk
```{r chunk9, include=FALSE}      
#NOTE: This code may take a little bit of time to run (~2 mins)

#relative risk estimation
relrisk_gail<-relative.risk(data_gail1, Raw_Ind=1)
#add relative risk to dataset
data_gail1<-cbind(data_gail1, relrisk_gail)
#Assign rel risk based on age <50 or >50
data_gail1$relrisk_gail<-ifelse(data_gail1$T1>=50, data_gail1$RR_Star2, data_gail1$RR_Star1)
#Calculate absolute risk
absrisk_gail<-absolute.risk(data_gail1, Raw_Ind=1, Avg_White=0)
#Add absolute risk to dataset
data_gail1$absrisk_gail<-absrisk_gail
#Change abs risk so it is not a % for our calibration functions
data_gail1$absrisk<-data_gail1$absrisk_gail/100

#Repeat for valdata2 (2-6y follow-up)
relrisk_gail<-relative.risk(data_gail2, Raw_Ind=1)
data_gail2<-cbind(data_gail2, relrisk_gail)
data_gail2$relrisk_gail<-ifelse(data_gail2$T1>=50, data_gail2$RR_Star2, data_gail2$RR_Star1)
absrisk_gail<-absolute.risk(data_gail2, Raw_Ind=1, Avg_White=0)
data_gail2$absrisk_gail<-absrisk_gail
data_gail2$absrisk<-data_gail2$absrisk_gail/100
```


```{r chunk10}
#Check the data: 
#Make sure mean for Error_Ind=0, otherwise there is an error in the file
check.summary(data_gail1, Raw_Ind=1, Avg_White=0)
check.summary(data_gail2, Raw_Ind=1, Avg_White=0)
```

```{r chunk11}
#OPTIONAL CODE: This will save your calculated risks only - replace "filepath here" with where you'd like to save
#This will be helpful to have on hand if you're willing to store the risks, but does not have to be done if preferred.

data_risks1<-data_gail1[,c("id","cohort","relrisk_gail","absrisk")]
data_risks2<-data_gail2[,c("id","cohort","relrisk_gail","absrisk")]

#Save absolute risks for Gail model for all if wanted- optional
#write.csv(data_gail1, "filepath/GailRisks_1to5y_CTS.csv")
#write.csv(data_gail2, "filepath/GailRisks_2to6y_CTS.csv")
```

## Relative and absolute risks 1-5y

```{r chunk12}
#Now we produce the summary estimates
head(data_gail1)
a<-summarizerisks(data_gail1, relrisk_gail)
b<-summarizerisks(data_gail1, absrisk_gail)
c<-summarizerisks(data_gail1, absrisk)
gailrisks1<-rbind(a,b,c)
rownames(gailrisks1)<-c("relrisk", "absriskpct","absrisk")
gailrisks1
```
## Relative and absolute risks 2-6y

```{r chunk13}
a<-summarizerisks(data_gail2, relrisk_gail)
b<-summarizerisks(data_gail2, absrisk_gail)
c<-summarizerisks(data_gail2, absrisk)
gailrisks2<-rbind(a,b,c)
rownames(gailrisks2)<-c("relrisk", "absriskpct","absrisk")
print(gailrisks2)
```

## Summary relative and absolute risks by race 1-5y

```{r chunk14}
a<-summarizerisks_race(data_gail1, relrisk_gail)
b<-summarizerisks_race(data_gail1, absrisk_gail)
gailrisks_byrace1<-as.data.frame(rbind(a,b))
gailrisks_byrace1
```
## Summary relative and absolute risks by race 2-6y

```{r chunk15}
a<-summarizerisks_race(data_gail2, relrisk_gail)
b<-summarizerisks_race(data_gail2, absrisk_gail)
gailrisks_byrace2<-as.data.frame(rbind(a,b))
gailrisks_byrace2
```

## Plot distributions

```{r chunk16,echo=FALSE}
#Plot distribution of relative risk in Gail model
ggplot(data_gail1, aes(x=relrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Relative Risk",y="Frequency") + ggtitle("Relative Risk for Gail Model 1-5y")
ggplot(data_gail2, aes(x=relrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Relative Risk",y="Frequency") + ggtitle("Relative Risk for Gail Model 2-6y")
```
```{r chunk17,echo=FALSE}
ggplot(data_gail1, aes(x=absrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Absolute Risk",y="Frequency") + ggtitle("Absolute Risk (%) for Gail Model 1-5y")
ggplot(data_gail2, aes(x=absrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Absolute Risk",y="Frequency") + ggtitle("Absolute Risk (%) for Gail Model 2-6y")
```


```{r chunk18, echo=FALSE, include=FALSE}
library(iCARE)
```


```{r chunk19, echo=FALSE}
#Now we will need to restructure dataset to allow for iCARE calibration 
## 5a. Change timevar to infinity if not a case (invasive or in situ). Print # invasive cases 


#Time of onset is time to disease for cases, otherwise Inf

data_gail1 <- data_gail1 %>%
  mutate(time.of.onset=ifelse(caseinv==1, timevar, Inf),
         sampling.fraction=ifelse(caseinv==1, 1, sampling.weights))

data_gail2 <- data_gail2 %>%
  mutate(time.of.onset=ifelse(caseinv==1, timevar, Inf),
         sampling.fraction=ifelse(caseinv==1, 1, sampling.weights))

## 5b. Add lp and subset

#ADD LP
data_gail1$lp<-log(data_gail1$relrisk_gail)
data_gail2$lp<-log(data_gail2$relrisk_gail)

#Create validation data so it matches iCARE 

#study.entry.age=age (same as T1)
#study.exit.age=age+timevar

data_gail1_val<-data_gail1[,c("id","T1","timevar","caseinv","relrisk_gail","absrisk","time.of.onset","sampling.fraction", "lp","Race", "cohort")]
data_gail2_val<-data_gail2[,c("id","T1","timevar","caseinv","relrisk_gail","absrisk","time.of.onset","sampling.fraction", "lp","Race", "cohort")] 

#set exit age : minimum of age+timediff and age+5
data_gail1_val$exitage<-pmin(data_gail1_val$T1+data_gail1_val$timevar, data_gail1_val$T1+5)
#If missing lastfup assume we have 5y
data_gail1_val$exitage<-ifelse(is.na(data_gail1_val$exitage),data_gail1_val$T1+5, data_gail1_val$exitage)

data_gail2_val$exitage<-pmin(data_gail2_val$T1+data_gail2_val$timevar, data_gail2_val$T1+5)
#If missing lastfup assume we have 5y
data_gail2_val$exitage<-ifelse(is.na(data_gail2_val$exitage),data_gail2_val$T1+5, data_gail2_val$exitage)

data_gail1_val<-data_gail1_val[,c("id","T1","exitage","caseinv","relrisk_gail","absrisk","time.of.onset","sampling.fraction", "lp","Race", "cohort")] 
data_gail2_val<-data_gail2_val[,c("id","T1","exitage","caseinv","relrisk_gail","absrisk","time.of.onset","sampling.fraction", "lp","Race", "cohort")] 

#Rename columns so it matches iCARE scheme
colnames(data_gail1_val)<-c("id","study.entry.age","study.exit.age","observed.outcome","relrisk_gail","absrisk", "time.of.onset","sampling.weights","lp", "race","cohort")
colnames(data_gail2_val)<-c("id","study.entry.age","study.exit.age","observed.outcome","relrisk_gail","absrisk", "time.of.onset","sampling.weights","lp", "race","cohort")
```

## Calibration Results

### E/O Overall

```{r chunk20,echo=FALSE}
#Gail overall: data_validategail_full
a<-obsexpected_overall(data_gail1_val)
a<-as.data.frame(a)
b<-obsexpected_overall(data_gail2_val)
b<-as.data.frame(b)
rownames(a)<-"1-5y"
rownames(b)<-"2-6y"
c<-rbind(a,b)
c
```

### E/O by decile

```{r chunk21,echo=FALSE}
#E/O by category of absolute risk
decile<-c(1:10)
eocat1<-obsexpected_bycat_abs(data_gail1_val)
eocat1<-cbind(decile, eocat1)
print("E/O by decile-1-5y")
eocat1
eocat2<-obsexpected_bycat_abs(data_gail2_val)
eocat2<-cbind(decile, eocat2)
print("E/O by decile-2-6y")
eocat2
```


### Absolute risk plot data

```{r chunk22, echo=FALSE}
absrisk_bycat_abs_gail1<-as.data.frame(returnabsrisk_abs(data_gail1_val))
print("Absolute risk by absolute risk decile")
absrisk_bycat_abs_gail1

absrisk_bycat_abs_gail2<-as.data.frame(returnabsrisk_abs(data_gail2_val))
print("Absolute risk by absolute risk decile")
absrisk_bycat_abs_gail2
```

### Calibration plot of absolute risk

```{r chunk23}
a1<-suppressWarnings(plotcalibration_abs(absrisk_bycat_abs_gail1))
plot(a1)

b1<-suppressWarnings(plotcalibration_abs(absrisk_bycat_abs_gail2))
plot(b1)
```

### Calibration plot of absolute risk #2

```{r chunk24}
a2<-suppressWarnings(plotcalibration_abs2(absrisk_bycat_abs_gail1))
plot(a2)

b2<-suppressWarnings(plotcalibration_abs2(absrisk_bycat_abs_gail2))
plot(b2)
```

## AUC for Gail model

### AUC for relative risk

```{r chunk25}
AUC1<-as.data.frame(AUCcalc(data_gail1_val))
AUC2<-as.data.frame(AUCcalc(data_gail1_val))
rownames(AUC1)<-"1-5y"
rownames(AUC2)<-"2-6y"
c<-rbind(AUC1, AUC2)
print("AUC")
c
```
### AUC absolute risk

```{r chunk26}
AUC1absrisk<-as.data.frame(AUCcalc_absrisk(data_gail1_val))

AUC2absrisk<-as.data.frame(AUCcalc_absrisk(data_gail2_val))
rownames(AUC1absrisk)<-"1-5y"
rownames(AUC2absrisk)<-"2-6y"
c<-rbind(AUC1absrisk, AUC2absrisk)
print("AUC absrisk")
c
```
## AUC absolute risk with age adjustment

```{r chunk27}
AUC1<-as.data.frame(AUCcalc_absrisk_ageadjust(data_gail1_val))
print("AUC age adjusted overall")
AUC1

AUC2<-as.data.frame(AUCcalc_absrisk_ageadjust(data_gail2_val))
print("AUC age adjusted overall")
AUC2

```


```{r instructions, include=FALSE}
#You will now need to run rmarkdown::render in your R Console. This cannot be included in the .Rmd file directly, or it will iteratively run.
#Please enter this code into the console:
#     rmarkdown::render("FILEPATH/Ext_Model_Calibration_Gail.Rmd")
#The resulting .html file will be called Ext_Model_Calibration_Gail as well, and will be found in the same file path as that of your .Rmd file.
```