---
title: "Ext_Model_Calibration_iCARELit"
output:
  html_document: default
    df_print: paged
date: "2025-04-15"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label="allow")
```

```{r include=FALSE}
#Install required packages and source functions

if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("dplyr"))
}
if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("tidyverse"))
}
if(!require(lubridate)){
  install.packages("lubridate")
  suppress.warnings(library("lubridate"))
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  suppress.warnings(library("ggplot2"))
}
if(!require(gtools)){
  install.packages("gtools")
  suppress.warnings(library("gtools"))
}

if(!require(BCRA)){
  install.packages("BCRA")
  suppress.warnings(library("BCRA"))
}


#Load the required functions
source("FILEPATH/Functions_ExistingModelValidation.R",local=knitr::knit_global())
```

```{r include=FALSE}

#set sampling fraction to 1 - needed variable in iCARE Lit
valdata$sampling.fraction<-1
valdata2$sampling.fraction<-1
```


```{r include=FALSE}

#This is the datasetup to calculate risks using iCARE-Lit.
#############
# 1 to 5y
#############
table(valdata$meno_age_cat5)
data_icare <- valdata %>%  
  #infinity if outside followup
  mutate(time.of.onset=ifelse(case_5yinv==1, timediff_case, Inf),
         sampling.fraction=ifelse(case_5yinv==1,1,
                                  sampling.fraction),
         
         height=height/10, 
         age=floor(age),
         agestart=age,
         ageend=pmin(age+5, age+timediff),
         age_at_menarche=relevel(as.factor(agemenarche_cat7), ref="4"),
         parity=as.factor(parity_cat),
         oc_current=ocuse_current,
         oc_ever=ocuse_ever,
         bbd=bbd_history,
         age_first_birth=ifelse(parity_new>0, ageflb_parous, 
                                ifelse(parity_new==0 | parous==0, 1, NA)),
         age_first_birth=as.factor(age_first_birth),
         famhist=famhx_first,
         alcohol_intake=as.factor(alcgm_cat7),
         age_at_menopause=ifelse(!is.na(meno_age_cat5), relevel(as.factor(meno_age_cat5), ref="4"),NA),
         hrt=as.factor(hrt_use_icare),
         hrt_type=pmh_type_icare)

```

```{r include=FALSE}
# Divide datasets for lt50 and ge50
# Note: If you have no, or very few participants <50 you will just run for the >=50 group.

data_icare_lt50<-data_icare[which (data_icare$age<50),]
data_icare_ge50<-data_icare[which (data_icare$age>=50),]

## Fit names to iCARE objects
## Note: bmi curc differs for lt50 and ge50

data_icare_lt50<- data_icare_lt50 %>% 
  mutate(bmi_curc=ifelse(bmi_cat4==1, 0, 
                         ifelse(bmi_cat4==2, 1, 
                                ifelse(bmi_cat4==3, 2, 
                                       ifelse(bmi_cat4==4,3, NA)))),
         bmi_curc=relevel(as.factor(bmi_curc), ref="1"))

data_icare_ge50<-data_icare_ge50 %>%
  mutate(bmi_curc=as.factor(bmi_cat3))

#Restrict for dataset 2 (2-6 y followup) so it doesn't go past 75 years as that is the limit.

data_icare_ge50<-data_icare_ge50[which (data_icare_ge50$age<76),]

data_icare_lt50_val<-data_icare_lt50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","sampling.fraction")]
#change names
names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.fraction")

data_icare_ge50_val<-data_icare_ge50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","sampling.fraction")]
#change names
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.fraction")

```

```{r include=FALSE}
## 2a. installation
if(!require(BiocManager)){
  install.packages("BiocManager")
  BiocManager::install("iCARE")
  suppress.warnings(library(iCARE))
}

# ** Data notes: iCARE dependencies include: *model.lit.lt50*, *cov_info_lit_lt50*, *model.lit.ge50*, and *cov_info_lit_ge50* apply to ALL studies**
# ** For US based studies, you can use *bc_inc_US* and *mort_US* and *ref_lit_lt50_US* and *ref_lit_ge50_US***
# ** If you are a non-US cohort, PLEASE USE reference datasets, incidence rates, and mortality rates that are relevant to your country/region**
# ** You will need to replace reference to datasets in 2b and 2c to align with your data **
# ** Variable names included in ref. dataset need to be same names as those in the validation dataset. Change names in validation dataset if needed.**

##  2b. <50 absolute risk

#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort
#Please note, these will take a bit to run.

results_iCARE_lt50=computeAbsoluteRisk(model.formula = model.lit.lt50, model.cov.info = cov_info_lit_lt50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_lt50,
                                       model.ref.dataset = ref_lit_lt50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_lt50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_lt50[, names(ref_lit_lt50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)

## 2c. >=50 absolute risk
#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort

results_iCARE_ge50=computeAbsoluteRisk(model.formula = model.lit.ge50, model.cov.info = cov_info_lit_ge50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_ge50,
                                       model.ref.dataset = ref_lit_ge50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_ge50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_ge50[, names(ref_lit_ge50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)

##Store risk estimates in dataframes for calibration

#store summary with abs risk as data frame so add in risk estimates

results_iCARE_lt50det<-as.data.frame(results_iCARE_lt50$details)
#add in lps
lps_lt50<-as.vector(results_iCARE_lt50$lps)
results_iCARE_lt50det$lps<-lps_lt50 

results_iCARE_ge50det<-as.data.frame(results_iCARE_ge50$details)
lps_ge50<-as.vector(results_iCARE_ge50$lps)
results_iCARE_ge50det$lps<-lps_ge50

#Dataset: results_iCARE_lt50det has Risk_Estimate- abs risk & lps 
#colnames(data_icare_lt50_val)
#Set absolute risk vector & lps
data_icare_lt50_val$absrisk<-results_iCARE_lt50det$Risk_Estimate
data_icare_lt50_val$lp<-results_iCARE_lt50det$lps

data_icare_ge50_val$absrisk<-results_iCARE_ge50det$Risk_Estimate
data_icare_ge50_val$lp<-results_iCARE_ge50det$lps

data_icare_lt50$absrisk<-results_iCARE_lt50det$Risk_Estimate
data_icare_ge50$absrisk<-results_iCARE_ge50det$Risk_Estimate

names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights",  "absrisk", "lp")
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights",  "absrisk", "lp")

# Combine datasets
data_icare_full<-rbind(data_icare_lt50_val, data_icare_ge50_val)
data_icare_full$absrisk<-as.numeric(data_icare_full$absrisk)

```

##Summary of iCARE-Lit risks 1-5y

```{r echo=FALSE}

a<-summarizerisks(data_icare_full, absrisk)
b<-summarizerisks(data_icare_full, lp)
icarerisks<-rbind(a,b)
rownames(icarerisks)<-c("AbsRisk","LP")
icarerisks
```

```{r include=FALSE}

#Save absolute risks and absolute risk cats for iCARE model for all if wanted
#data_risks1<-data_icare_full[,c("id","cohort","absrisk","lp")]
#write.csv(data_risks1, "filepath/icarerisks_1to5y_CPS2.csv") #rename if desired

```
##Model calibration 1-5y

###E/O overall

```{r echo=FALSE}

## E/O invasive
eo_invas<-as.data.frame(obsexpected_overall(data_icare_full))
print("E/O invasive only")
eo_invas
```

###E/O by decile 1-5y

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_icare_full))
print("E/O by cat, invasive")
eo_bycat_invas
```

##AUCs for iCARE Lit model

###AUC by relative risk 1-5y

```{r echo=F}
AUC1<-as.data.frame(AUCcalc(data_icare_full))
print("AUC by relative risk")
AUC1
```
##AUC absolute risk 1-5y

```{r echo=FALSE}
AUCageinc<-as.data.frame(AUCcalc_absrisk(data_icare_full))
print("AUC age incorporated")
AUCageinc
```

##AUC absolute risk age adjusted 1-5y

```{r echo=FALSE}
AUCageinc<-as.data.frame(AUCcalc_absrisk_ageadjust(data_icare_full))
print("AUC age incorporated")
AUCageinc

```

## Calibration plots and data 1-5y

```{r}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_icare_full))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs

a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

```{r include=FALSE} 
#This repeats all done above for the follow up period 2-6 y (valdata2)
#2-6y

 data_icare <- valdata2 %>%  
   #infinity if outside followup
   mutate(time.of.onset=ifelse(case_6yinv==1, timediff_case, Inf),
          sampling.fraction=ifelse(case_6yinv==1,1,
                                   sampling.fraction),
          
          height=height/10, 
          age=floor(age),
          agestart=age,
          ageend=pmin(age+5, age+timediff),
          age_at_menarche=relevel(as.factor(agemenarche_cat7), ref="4"),
          parity=as.factor(parity_cat),
          oc_current=ocuse_current,
          oc_ever=ocuse_ever,
          bbd=bbd_history,
          age_first_birth=ifelse(parity_new>0, ageflb_parous, 
                                 ifelse(parity_new==0 | parous==0, 1, NA)),
          age_first_birth=as.factor(age_first_birth),
          famhist=famhx_first,
          alcohol_intake=as.factor(alcgm_cat7),
          age_at_menopause=ifelse(!is.na(meno_age_cat5), relevel(as.factor(meno_age_cat5), ref="4"), NA),
          hrt=as.factor(hrt_use_icare),
          hrt_type=pmh_type_icare)
 
 ###########################
 # Divide for lt50 and ge50
 ###########################
 

data_icare_lt50<-data_icare[which (data_icare$age<50),]

data_icare_ge50<-data_icare[which (data_icare$age>=50),]

##  1b. Fit names to iCARE objects
## bmi curc differs for lt50 and ge50
#can't relevel b/c ref isn't in dataset.
data_icare_lt50<- data_icare_lt50 %>% 
  mutate(bmi_curc=ifelse(bmi_cat4==1, 0, 
                         ifelse(bmi_cat4==2, 1, 
                                ifelse(bmi_cat4==3, 2, 
                                       ifelse(bmi_cat4==4,3, NA)))),
         bmi_curc=relevel(as.factor(bmi_curc), ref="1"))

data_icare_ge50<-data_icare_ge50 %>%
  mutate(bmi_curc=as.factor(bmi_cat3))

#Restrict for #2 so it doesn't go past 75 years

data_icare_ge50<-data_icare_ge50[which (data_icare_ge50$age<76),]


data_icare_lt50_val<-data_icare_lt50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","sampling.fraction")]
#change names
names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.fraction")

data_icare_ge50_val<-data_icare_ge50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","sampling.fraction")]
#change names
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.fraction")

## 2a. installation

#install.packages("BiocManager")
#BiocManager::install("iCARE")
library(iCARE)

# ** Data notes: iCARE dependencies include: *model.lit.lt50*, *cov_info_lit_lt50*, *model.lit.ge50*, and *cov_info_lit_ge50* apply to ALL studies**
# ** For US based studies, you can use *bc_inc_US* and *mort_US* and *ref_lit_lt50_US* and *ref_lit_ge50_US***
# ** If you are a non-US cohort, PLEASE USE reference datasets, incidence rates, and mortality rates that are relevant to your country/region**
# ** You will need to replace reference to datasets in 2b and 2c to align with your data **
# ** Variable names included in ref. dataset need to be same names as those in the validation dataset. Change names in validation dataset if needed.**

##  2b. <50 absolute risk

#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort


results_iCARE_lt50=computeAbsoluteRisk(model.formula = model.lit.lt50, model.cov.info = cov_info_lit_lt50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_lt50,
                                       model.ref.dataset = ref_lit_lt50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_lt50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_lt50[, names(ref_lit_lt50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)



## 2c. >=50 absolute risk
#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort

results_iCARE_ge50=computeAbsoluteRisk(model.formula = model.lit.ge50, model.cov.info = cov_info_lit_ge50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_ge50,
                                       model.ref.dataset = ref_lit_ge50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_ge50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_ge50[, names(ref_lit_ge50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)

##2d. Store risk estimates in dataframes for calibration

#store summary with abs risk as data frame so add in risk estimates

results_iCARE_lt50det<-as.data.frame(results_iCARE_lt50$details)
#add in lps
lps_lt50<-as.vector(results_iCARE_lt50$lps)
results_iCARE_lt50det$lps<-lps_lt50 

results_iCARE_ge50det<-as.data.frame(results_iCARE_ge50$details)
lps_ge50<-as.vector(results_iCARE_ge50$lps)
results_iCARE_ge50det$lps<-lps_ge50

#Dataset: results_iCARE_lt50det has Risk_Estimate- abs risk & lps 
#colnames(data_icare_lt50_val)
#Set absolute risk vector & lps
data_icare_lt50_val$absrisk<-results_iCARE_lt50det$Risk_Estimate
data_icare_lt50_val$lp<-results_iCARE_lt50det$lps

data_icare_ge50_val$absrisk<-results_iCARE_ge50det$Risk_Estimate
data_icare_ge50_val$lp<-results_iCARE_ge50det$lps

data_icare_lt50$absrisk<-results_iCARE_lt50det$Risk_Estimate
data_icare_ge50$absrisk<-results_iCARE_ge50det$Risk_Estimate

names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights",  "absrisk", "lp")
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights",  "absrisk", "lp")
data_icare_full<-rbind(data_icare_lt50_val, data_icare_ge50_val)
data_icare_full$absrisk<-as.numeric(data_icare_full$absrisk)
```

##Summary of iCARE risks 2-6y
```{r echo=FALSE}
# Step 3. Combine datasets
a<-summarizerisks(data_icare_full, absrisk)
b<-summarizerisks(data_icare_full, lp)
icarerisks<-rbind(a,b)
rownames(icarerisks)<-c("AbsRisk","LP")
icarerisks
```
```{r include=FALSE}
#Optional

 #Save absolute risks and absolute risk cats for icare model for all if wanted
# data_risks2<-data_icare_full[,c("id","cohort","absrisk","lp")]
# write.csv(data_risks2, "filepath/icare_2to6y_risks_CTS.csv") #rename if you want

```
##Calibration for 2-6y

###E/O overall 2-6 y

```{r echo=FALSE}
 # Step 4. Calibration

 ## 4a. E/O invasive
 eo_invas<-as.data.frame(obsexpected_overall(data_icare_full))
 print("E/O invasive only")
 eo_invas
```

###E/O by decile 2-6 y

```{r echo=FALSE} 
 eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_icare_full))
 print("E/O by cat, invasive")
 eo_bycat_invas
```

##AUC for iCARE risks, 2-6y

###AUC by rel risk 2-6y

```{r echo=F}
AUC1<-as.data.frame(AUCcalc(data_icare_full))
print("AUC by rel risk")
AUC1
```
##AUC absolute risk 2-6y

```{r echo=FALSE}
AUCageinc<-as.data.frame(AUCcalc_absrisk(data_icare_full))
print("AUC by absolute risk")
AUCageinc
```

##AUC absolute risk age adjusted 2-6y

```{r echo=FALSE}
AUCageinc<-as.data.frame(AUCcalc_absrisk_ageadjust(data_icare_full))
print("AUC by absolute risk, age-adjusted")
AUCageinc

```
 
## Calibration plots

```{r echo=FALSE} 
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_icare_full))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs
 
 
 ## Calibration plot of absolute risk
 
 a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
 plot(a1)
 ## Calibration plot of absolute risk #2
```

```{r instructions, include=FALSE}
#You will now need to run rmarkdown::render in your R Console. This cannot be included in the .Rmd file directly, or it will iteratively run.
#Please enter this code into the console:
#     rmarkdown::render("FILEPATH/Ext_Model_Calibration_iCARELit.Rmd")
#The resulting .html file will be called Ext_Model_Calibration_Gail as well, and will be found in the same file path as that of your .Rmd file.
```

