---
title: "Ext_Model_Calibration_TC.Rmd"
output: html_document
date: "2025-04-16"
---

```{r include=FALSE}
#Clear r environment
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label="allow")
```

```{r chunk1, echo=FALSE, include=FALSE}
#Here you will load all required packages and functions
#functions
source("FILEPATH/Functions_ExistingModelValidation.R",local=knitr::knit_global())

if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("dplyr"))
}
if(!require(tidyverse)){
  install.packages("tidyverse")
  suppress.warnings(library("tidyverse"))
}
if(!require(lubridate)){
  install.packages("lubridate")
  suppress.warnings(library("lubridate"))
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  suppress.warnings(library("ggplot2"))
}
if(!require(gtools)){
  install.packages("gtools")
  suppress.warnings(library("gtools"))
}
```

```{r include=FALSE}
#Step 1. Create file in proper format.

#Read in risk datasets created previously
#Datasets from 1-5y evaluation, assumption 1
#You may have a different number of datasets- it depends on cohort size.

#Assumption 1

#1 to 5y
TCrisk<-read.delim("FILEPATH/data_1to5y_assump1_risks1",header=FALSE, sep="\t")
TCrisk2<-read.delim("FILEPATH/data_1to5y_assump1_risks2",header=FALSE, sep="\t")
TCrisk3<-read.delim("FILEPATH/data_1to5y_assump1_risks3",header=FALSE, sep="\t")
TCrisk4<-read.delim("FILEPATH/data_1to5y_assump1_risks4",header=FALSE, sep="\t")
TCrisk5<-read.delim("FILEPATH/data_1to5y_assump1_risks5",header=FALSE, sep="\t")

TCrisk_assump1_1to5y<-rbind(TCrisk,TCrisk2, TCrisk3, TCrisk4, TCrisk5)
TCrisk_assump1_1to5y<-TCrisk_assump1_1to5y[,c(1,3)]
names(TCrisk_assump1_1to5y)<-c("tyid","absrisk")


#2 to 6y
TCrisk<-read.delim("FILEPATH/data_2to6y_assump1_risks1",header=FALSE, sep="\t")
TCrisk2<-read.delim("FILEPATH/data_2to6y_assump1_risks2",header=FALSE, sep="\t")
TCrisk3<-read.delim("FILEPATH/data_2to6y_assump1_risks3",header=FALSE, sep="\t")
TCrisk4<-read.delim("FILEPATH/data_2to6y_assump1_risks4",header=FALSE, sep="\t")
TCrisk5<-read.delim("FILEPATH/data_2to6y_assump1_risks5",header=FALSE, sep="\t")

TCrisk_assump1_2to6y<-rbind(TCrisk,TCrisk2, TCrisk3, TCrisk4, TCrisk5)
TCrisk_assump1_2to6y<-TCrisk_assump1_2to6y[,c(1,3)]
names(TCrisk_assump1_2to6y)<-c("tyid","absrisk")


#Assumption 2

#1 to 5y
TCrisk<-read.delim("FILEPATH/data_1to5y_assump2_risks1",header=FALSE, sep="\t")
TCrisk2<-read.delim("FILEPATH/data_1to5y_assump2_risks2",header=FALSE, sep="\t")
TCrisk3<-read.delim("FILEPATH/data_1to5y_assump2_risks3",header=FALSE, sep="\t")
TCrisk4<-read.delim("FILEPATH/data_1to5y_assump2_risks4",header=FALSE, sep="\t")
TCrisk5<-read.delim("FILEPATH/data_1to5y_assump2_risks5",header=FALSE, sep="\t")

TCrisk_assump2_1to5y<-rbind(TCrisk,TCrisk2, TCrisk3, TCrisk4, TCrisk5)
TCrisk_assump2_1to5y<-TCrisk_assump2_1to5y[,c(1,3)]
names(TCrisk_assump2_1to5y)<-c("tyid","absrisk")

#2 to 6y
TCrisk<-read.delim("FILEPATH/data_2to6y_assump2_risks1",header=FALSE, sep="\t")
TCrisk2<-read.delim("FILEPATH/data_2to6y_assump2_risks2",header=FALSE, sep="\t")
TCrisk3<-read.delim("FILEPATH/data_2to6y_assump2_risks3",header=FALSE, sep="\t")
TCrisk4<-read.delim("FILEPATH/data_2to6y_assump2_risks4",header=FALSE, sep="\t")
TCrisk5<-read.delim("FILEPATH/data_2to6y_assump2_risks5",header=FALSE, sep="\t")

TCrisk_assump2_2to6y<-rbind(TCrisk,TCrisk2, TCrisk3, TCrisk4, TCrisk5)
TCrisk_assump2_2to6y<-TCrisk_assump2_2to6y[,c(1,3)]
names(TCrisk_assump2_2to6y)<-c("tyid","absrisk")
```

##Risk summaries for all datasets

```{r echo=FALSE}
risksummary1<-summarizerisks(TCrisk_assump1_1to5y,absrisk)
risksummary2<-summarizerisks(TCrisk_assump1_2to6y,absrisk)
risksummary3<-summarizerisks(TCrisk_assump2_1to5y,absrisk)
risksummary4<-summarizerisks(TCrisk_assump2_2to6y,absrisk)

risks<-rbind(risksummary1, risksummary2, risksummary3, risksummary4)
risks
```


```{r include=FALSE}

################
# Combine files
################

#Link absolute risk data with TC data by merging all 
#You will need to run Step2.Create2datasets.Rmd so that you have the datasets named "valdata" and "valdata2" stored in your R environment.

head(valdata2$subject_id)
head(TCrisk_assump2_1to5y)

valdata<-valdata %>%
  mutate(tyid=gsub("_","",subject_id),
         tyid=as.character(tyid),
         sampling.fraction=1)

valdata2<-valdata2 %>%
  mutate(tyid=gsub("_","",subject_id),
         tyid=as.character(tyid),
         sampling.fraction=1)


TCdata<-valdata[,c("tyid","race","age","timediff","timediff_case","case_5yinv","cohort","sampling.fraction")]
TCdata<-merge(TCdata, TCrisk_assump1_1to5y, by="tyid")

TCdata2<-valdata2[,c("tyid","race","age","timediff","timediff_case","case_6yinv","cohort","sampling.fraction")]
TCdata2<-merge(TCdata2, TCrisk_assump1_2to6y, by="tyid")

TCdata3<-valdata[,c("tyid","race","age","timediff","timediff_case","case_5yinv","cohort","sampling.fraction")]
TCdata3<-merge(TCdata3, TCrisk_assump2_1to5y, by="tyid")

TCdata4<-valdata2[,c("tyid","race","age","timediff","timediff_case","case_6yinv","cohort","sampling.fraction")]
TCdata4<-merge(TCdata4, TCrisk_assump2_2to6y, by="tyid")

#Note: If there is an issue with merging, it is likely the tyid variable. Make sure the format of "tyid" and your original "id" variable are the same. See the code above where gsub is used to transform the id variable and change if needed. 
```

```{r include=FALSE}
#############################
# Calibration with iCARE    #
#############################
if(!require(BiocManager)){
  install.packages("BiocManager")
  BiocManager::install("iCARE")
  suppress.warnings(library(iCARE))
}

data_TCval<-TCdata %>%
  mutate(time.of.onset=ifelse(case_5yinv==1, timediff_case, Inf),
         sampling.fraction=ifelse(case_5yinv==1,1,
                                  sampling.fraction),
         ageend=pmin(age+5, age+timediff))

#Keep needed columns
data_TCval<-data_TCval[,c("tyid","race","cohort","age","ageend","case_5yinv","absrisk","time.of.onset","sampling.fraction")]
#Rename columns so it matches iCARE scheme
colnames(data_TCval)<-c("tyid","race","cohort","study.entry.age","study.exit.age","observed.outcome","absrisk", "time.of.onset","sampling.weights")



data_TCval2<-TCdata2 %>%
  mutate(time.of.onset=ifelse(case_6yinv==1, timediff_case, Inf),
         sampling.fraction=ifelse(case_6yinv==1,1,
                                  sampling.fraction),
         ageend=pmin(age+5, age+timediff))
#Keep needed columns

data_TCval2<-data_TCval2[,c("tyid","race","cohort","age","ageend","case_6yinv","absrisk","time.of.onset","sampling.fraction")]
#Rename columns so it matches iCARE scheme
colnames(data_TCval2)<-c("tyid","race","cohort","study.entry.age","study.exit.age","observed.outcome","absrisk", "time.of.onset","sampling.weights")



############
#assumption 2

data_TCval3<-TCdata3 %>%
  mutate(time.of.onset=ifelse(case_5yinv==1, timediff_case, Inf),
         sampling.fraction=ifelse(case_5yinv==1,1,
                                  sampling.fraction),
         ageend=pmin(age+5, age+timediff))
#Keep needed columns

data_TCval3<-data_TCval3[,c("tyid","race","cohort","age","ageend","case_5yinv","absrisk","time.of.onset","sampling.fraction")]
#Rename columns so it matches iCARE scheme
colnames(data_TCval3)<-c("tyid","race","cohort","study.entry.age","study.exit.age","observed.outcome","absrisk", "time.of.onset","sampling.weights")


data_TCval4<-TCdata4 %>%
  mutate(time.of.onset=ifelse(case_6yinv==1, timediff_case, Inf),
         sampling.fraction=ifelse(case_6yinv==1,1,
                                  sampling.fraction),
         ageend=pmin(age+5, age+timediff))
#Keep needed columns

data_TCval4<-data_TCval4[,c("tyid","race","cohort","age","ageend","case_6yinv","absrisk","time.of.onset","sampling.fraction")]
#Rename columns so it matches iCARE scheme
colnames(data_TCval4)<-c("tyid","race","cohort","study.entry.age","study.exit.age","observed.outcome","absrisk", "time.of.onset","sampling.weights")
```
# Assumption 1, 1-5y

## E/O overall

```{r echo=FALSE}

## 4a. E/O invasive
eo_invas<-as.data.frame(obsexpected_overall(data_TCval))
print("E/O invasive only")
eo_invas
```

## E/O by decile

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_TCval))
print("E/O by cat, invasive")
eo_bycat_invas
```

## AUC by relative risk

```{r echo=FALSE}
AUC<-as.data.frame(AUCcalc(data_TCval))
print("AUC rel risk")
AUCabsrisk
```

## AUC by absolute risk

```{r echo=FALSE}
AUCabsrisk<-as.data.frame(AUCcalc_absrisk(data_TCval))
print("AUC absolute risk")
AUCabsrisk
```

## AUC by absolute risk age adjusted

```{r echo=FALSE}
AUCabsrisk_ageadjust<-as.data.frame(AUCcalc_absrisk_ageadjust(data_TCval))
print("AUC absolute risk, age adjust")
AUCabsrisk_ageadjust
```

## Plots
```{r echo=FALSE}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_TCval))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs

## Calibration plot of absolute risk
library(ggplot2)
a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

# Assumption 1, 2-6y

## E/O overall

```{r echo=FALSE}

## 4a. E/O invasive
eo_invas<-as.data.frame(obsexpected_overall(data_TCval2))
print("E/O invasive only")
eo_invas
```

## E/O by decile

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_TCval2))
print("E/O by cat, invasive")
eo_bycat_invas
```

## AUC by relative risk

```{r echo=FALSE}
AUC<-as.data.frame(AUCcalc(data_TCval2))
print("AUC rel risk")
AUCabsrisk
```

## AUC by absolute risk

```{r echo=FALSE}
AUCabsrisk<-as.data.frame(AUCcalc_absrisk(data_TCval2))
print("AUC absolute risk")
AUCabsrisk
```

## AUC by absolute risk age adjusted

```{r echo=FALSE}
AUCabsrisk_ageadjust<-as.data.frame(AUCcalc_absrisk_ageadjust(data_TCval2))
print("AUC absolute risk, age adjust")
AUCabsrisk_ageadjust
```

## Plots
```{r echo=FALSE}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_TCval2))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs

## Calibration plot of absolute risk
library(ggplot2)
a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

#Assumption 2, 1-5y

## E/O overall

```{r echo=FALSE}

## 4a. E/O invasive
eo_invas<-as.data.frame(obsexpected_overall(data_TCval3))
print("E/O invasive only")
eo_invas
```

## E/O by decile

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_TCval3))
print("E/O by cat, invasive")
eo_bycat_invas
```

## AUC by relative risk

```{r echo=FALSE}
AUC<-as.data.frame(AUCcalc(data_TCval3))
print("AUC rel risk")
AUCabsrisk
```

## AUC by absolute risk

```{r echo=FALSE}
AUCabsrisk<-as.data.frame(AUCcalc_absrisk(data_TCval3))
print("AUC absolute risk")
AUCabsrisk
```

## AUC by absolute risk age adjusted

```{r echo=FALSE}
AUCabsrisk_ageadjust<-as.data.frame(AUCcalc_absrisk_ageadjust(data_TCval3))
print("AUC absolute risk, age adjust")
AUCabsrisk_ageadjust
```

## Plots
```{r echo=FALSE}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_TCval3))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs

## Calibration plot of absolute risk
library(ggplot2)
a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

#Assumption 2, 2-6y

```{r echo=FALSE}

## 4a. E/O invasive
eo_invas<-as.data.frame(obsexpected_overall(data_TCval4))
print("E/O invasive only")
eo_invas
```

## E/O by decile

```{r echo=FALSE}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_TCval4))
print("E/O by cat, invasive")
eo_bycat_invas
```

## AUC by relative risk

```{r echo=FALSE}
AUC<-as.data.frame(AUCcalc(data_TCval4))
print("AUC rel risk")
AUCabsrisk
```

## AUC by absolute risk

```{r echo=FALSE}
AUCabsrisk<-as.data.frame(AUCcalc_absrisk(data_TCval4))
print("AUC absolute risk")
AUCabsrisk
```

## AUC by absolute risk age adjusted

```{r echo=FALSE}
AUCabsrisk_ageadjust<-as.data.frame(AUCcalc_absrisk_ageadjust(data_TCval4))
print("AUC absolute risk, age adjust")
AUCabsrisk_ageadjust
```

## Plots
```{r echo=FALSE}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_TCval4))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs

## Calibration plot of absolute risk
library(ggplot2)
a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

```{r instructions, include=FALSE}
#You will now need to run rmarkdown::render in your R Console. This cannot be included in the .Rmd file directly, or it will iteratively run.
#Please enter this code into the console:
#     rmarkdown::render("FILEPATH/Ext_Model_Calibration_TyrerCuzick.Rmd")
#The resulting .html file will be called Ext_Model_Calibration_Gail as well, and will be found in the same file path as that of your .Rmd file.
```
