```{r include=FALSE}
#Instructions:

# This file will read in your clean validation dataset, created after
# running "Step1.DataCleaning.R" and will perform model calibration
# using iCARE-Validate for the Gail (BCRAT) model.

# PLEASE INSERT COHORT ACRONYM IN [] BELOW for the FILENAME
# AND INSERT RUN DATE

# Run Notes: You will need to install these packages if you don't have them (available online):
# #BCRAT packages - install.packages("BCRA") - please run outside of Rmarkdown:
#  https://dceg.cancer.gov/tools/risk-assessment/bcra
# #iCARE packages - install.packages("BiocManager")
  #                  BiocManager::install("iCARE")
# https://www.bioconductor.org/packages/release/bioc/html/iCARE.html

# After specifying your filepaths, (ctrl+f: "filepath" for where to insert), you can simply:
# hit "Knit -> Knit to Word" and this will run the full document and produce output.

# Please send the finalized document to kristen_brantley@dfci.harvard.edu
# If any issues arise when trying to run this document, please reach out to Kristen Brantley. 
```


---
title: "Ext_Model_Calibration_Gail [COHORT ACRONYM]"
output: html_document
date: "MM-DD-YYYY"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Dataset: This is the dataset created from Step1.DataCleaning.R 
#Please ensure you use the correct data file, as naming conventions need to match.

valdata<-read.csv("filepath/bcrpp_valdata.R")
```

```{r}
#Functions required
source("FILEPATH/Functions_ExistingModelValidation.R", local=knitr::knit_global())
```

# Step 1. Load required packages

```{r echo=FALSE, include=FALSE}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('dplyr'))
suppressWarnings(library("lubridate"))
suppressWarnings(library("gtools"))
suppressWarnings(library("ggplot2"))

suppressWarnings(library(BCRA))
data("BrCa_lambda1")
data("BrCa_lambda2")
data("BrCa_beta")
```
          Note: If you need to explore the data set up further, 
          this will read in the example data from BCRAT tool (remove #)
          
          #```{r}
          
          #data(exampledata, package="BCRA")
          #mean(exampledata$RR_Star1)
          #check.summary(exampledata)
          #```
          
# Step 2. In Gail model, all missing variables are set to the most common

## 2a. Create variables for Gail model
```{r echo = T, results = 'hide'}
data_gail <-valdata[,c("subject_id","id","age","race","ethnicity","agemenarche","ageflb","famhx_first","biopsy_num","atyp_hyp","parous","cohort", "case_5yinv", "case_5y", "timediff_case_yr")]
```
## 2b. Reassign values based on Gail parameters
```{r include=FALSE}       
        # Missing data assignments
        # The standards used in the online Gail model were used for variables with missing data. 
        # Women with unknown age at menarche were assigned menarche at 14 years of age or older. 
        # Women with unknown age at first live birth were classified as giving birth before age 20 years. 
        # Women with missing family history were classified as having no family history. 
        # To verify agreement between the code we used and the online tool, we randomly selected 10 patients 
        # and compared the 5-year and lifetime risk estimates obtained from the code given us to those from the
        # online risk assessment tool. 
        # 
        # For our purposes, assign all Asian under "Chinese" for abs.risk calibration
        # Consider all hispanics american born
        # 1=Wh White 1983-87 SEER rates (rates used in NCI BCRAT)
        # 2=AA African-American
        # 3=HU Hispanic-American (US born) 1995-04
        # 4=NA Other (Native American and unknown race)
        # 5=HF Hispanic-American (Foreign born) 1995-04
        # 6=Ch Chinese-American
        # 7=Ja Japanese-American
        # 8=Fi Filipino-American
        # 9=Hw Hawaiian-American
        # 10=oP Other Pacific Islander
        # 11=oA Other Asian

       # Set end age as age+5 and +10 years
```

```{r}
data_gail <- data_gail %>%
  mutate(gailrace=ifelse(race==1 & ethnicity==1, 3, ifelse(race==1 & ethnicity !=1 ,1, ifelse(race==2, 2, ifelse(race==5, 4, ifelse(race==3, 6,    ifelse(race==4, 9, 4)))))),
         gailrace=replace(gailrace, is.na(gailrace), 4),
         gailagemen=ifelse(is.na(agemenarche), 99, agemenarche),
         gailageflb=replace(ageflb, is.na(ageflb) & parous==0, 98),
         gailageflb=replace(gailageflb, is.na(gailageflb), 99),
         gailnrels=ifelse(is.na(famhx_first), 99, famhx_first),
         gailbiop=ifelse(is.na(biopsy_num), 99, biopsy_num),
         gailatyp=ifelse(is.na(atyp_hyp), 99, atyp_hyp),
         gailageend5y=age+5, 
         gailageend10y=age+10)
```

## 2c. Keep only gail defined variables 
```{r echo = T, results = 'hide'}
#5y risk
data_gail<-data_gail[,c("subject_id","cohort","age","gailageend5y","gailrace","gailagemen","gailageflb","gailnrels","gailbiop","gailatyp", "case_5yinv", "case_5y", "timediff_case_yr")]
#Change names
names(data_gail)<-c("id","cohort","T1","T2","Race","AgeMen","Age1st","N_Rels","N_Biop","HypPlas","caseinv", "case_any", "timevar")
```


## Data check! 
      Make sure the variables are recoded correctly and check errors
```{r echo = T}
      recodechecktable<-recode.check(data_gail, Raw_Ind=1)
      test<-recodechecktable[which (recodechecktable$Error_Ind>0),]
      test
      #If any observations show up on this print out then these have errors 
      #that need to be fixed in the assignment of raw gail vars.
      #If so, please Contact Kristen Brantley
```

# Step 3. Estimate relative and absolute risks from Gail model

## 3a.Estimate Relative risk
```{r}      
#Estimate rel risk
relrisk_gail<-relative.risk(data_gail, Raw_Ind=1)
    #check.summary(relrisk_gail)
data_gail<-cbind(data_gail, relrisk_gail)
```

```{r}
#Assign rel risk based on age <50 or >50
data_gail$relrisk_gail<-ifelse(data_gail$T1>=50, data_gail$RR_Star2, data_gail$RR_Star1)
```
## 3b. Calculate Absolute risk
```{r}
    absrisk_gail<-absolute.risk(data_gail, Raw_Ind=1, Avg_White=0)
    #check.summary(data_gail, Raw_Ind=1, Avg_White=0)
```

```{r}
#Assign absrisk in dataset
data_gail$absrisk_gail<-absrisk_gail
```

### Data check!
Make sure mean for Error_Ind=0, otherwise there is an error in the file
Note, abs risk is given as a % for 5 years
```{r}
check.summary(data_gail, Raw_Ind=1, Avg_White=0)
```

## 3c. Classify in groups
For comparison, classify individuals into relative risk categories & absolute risk categories (tertiles)
<15%, 15-20%, >20%
```{r}
data_gail$relriskcat<-ifelse(data_gail$relrisk_gail<1.15, 1, ifelse(data_gail$relrisk_gail>=1.15 & data_gail$relrisk_gail<1.54, 2, 3))
data_gail$absrisk_gail<-as.numeric(absrisk_gail)
data_gail$absriskcat<-quantcut(data_gail$absrisk_gail, q=3)
#ties are in same group
data_gail$absriskcat_num<-as.numeric(data_gail$absriskcat)
data_gail$absrisk<-data_gail$absrisk_gail/100
```

```{r}
#OPTIONAL
#Save absolute risks and absolute risk cats for Gail model for all if wanted- optional
write.csv(data_gail, "filepath/Gail Model/Gail_5yrisks.csv")
```

# Step 4.Descriptive data for Gail risks

## 4a. Summary relative and absolute risks overall
```{r}
a<-summarizerisks(data_gail, relrisk_gail)
b<-summarizerisks(data_gail, absrisk_gail)
c<-summarizerisks(data_gail, absrisk)
gailrisks1<-rbind(a,b,c)
rownames(gailrisks1)<-c("relrisk", "absriskpct","absrisk")
gailrisks1
```

## 4b. Summary relative and absolute risks by race
```{r}
a<-summarizerisks_race(data_gail, relrisk_gail)
b<-summarizerisks_race(data_gail, absrisk_gail)
gailrisks_byrace<-as.data.frame(rbind(a,b))
gailrisks_byrace
```

## 4c. Plot distributions

```{r}
#Plot distribution of relative risk in Gail model
ggplot(data_gail, aes(x=relrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Relative Risk",y="Frequency") + ggtitle("Relative Risk for Gail Model")
```
```{r}
ggplot(data_gail, aes(x=absrisk_gail))+ geom_histogram(binwidth=0.5, fill="lightblue")+labs(x="Absolute Risk",y="Frequency") + ggtitle("Absolute Risk (%) for Gail Model")
   
```


# Step 5. iCARE Validate to test Calibration


```{r echo=FALSE, include=FALSE}
library(iCARE)
```

## 5a. Change timevar to infinity if not a case (invasive or in situ). Print # invasive cases 

```{r}
data_gail$time.of.onset<-ifelse(data_gail$caseinv==1, data_gail$timevar, Inf)
data_gail$time.of.onset2<-ifelse(data_gail$case_any==1, data_gail$timevar, Inf)

table(data_gail$caseinv,data_gail$case_any)

```


## 5b. Add lp and subset
```{r}
#ADD LP
data_gail$lp<-log(data_gail$relrisk_gail)

#E/O

data_gail$sampling.weights<-1
data_gail_validate_inv<-data_gail[,c("id","T1","T2","caseinv","relrisk_gail","absrisk","time.of.onset","sampling.weights", "lp","Race", "cohort")] 

#Rename columns so it matches iCARE scheme
colnames(data_gail_validate_inv)<-c("id","study.entry.age","study.exit.age","observed.outcome","relrisk_gail","absrisk", "time.of.onset","sampling.weights","lp", "race","cohort")
```



# Step 6. E/O Calculations

## 6a. E/O Overall
```{r}
#Gail overall: data_validategail_full
a<-obsexpected_overall(data_gail_validate_inv)
a<-as.data.frame(a)
rownames(a)<-"All"
print("E/O overall-Gail")
a
```

## 6b. E/O by decile
```{r}
#E/O by category of absolute risk
decile<-c(1:10)
eocat<-obsexpected_bycat_abs(data_gail_validate_inv)
eocat<-cbind(decile, eocat)
print("E/O by decile-Gail")
eocat
```

# Step 7. Calibration Plots

## 7a.Create plot data

```{r}
absrisk_bycat_abs_gail<-as.data.frame(returnabsrisk_abs(data_gail_validate_inv))
print("Absolute risk by absolute risk decile")
absrisk_bycat_abs_gail
```

## Calibration plot of absolute risk
```{r}
a1<-suppressWarnings(plotcalibration_abs(absrisk_bycat_abs_gail))
plot(a1)
```

## Calibration plot of absolute risk #2
```{r}
a1b<-suppressWarnings(plotcalibration_abs2(absrisk_bycat_abs_gail))
plot(a1b)
```

# AUC for Gail model
```{r echo=F}
AUC<-as.data.frame(AUCcalc(data_gail_validate_inv))
print("AUC overall")
AUC
```

```{r echo=F}
AUCageinc<-as.data.frame(AUCcalc_ageinc(data_gail_validate_inv))
print("AUC age incorporated")
AUCageinc
```
