```{r include=FALSE}
# This file will read in your clean validation dataset, created after
# running "Step1.DataCleaning.R" and will perform model calibration
# using iCARE-Validate for the iCARE model.

# Run Notes: Please insert cohort acronym in title and the date of run.

# After specifying your filepaths, (ctrl+f: "filepath" for where to insert), you can simply:
# hit "Knit -> Knit to Word" and this will run the full document and produce output.

# Please send the finalized document to kristen_brantley@dfci.harvard.edu
# If any issues arise when trying to run this document, please reach out to Kristen Brantley. 

# Before running iCARE you will need to load the iCARE dependencies to your R environment.
# These are contained in the ".zip" file- labeled "icare-lit-r-dependencies" in GitHub

#PLEASE INSERT COHORT ACRONYM IN [] BELOW for the FILENAME
#AND INSERT RUN DATE
```

---
title: "Model Validation iCARE [COHORT ACRONYM]"
output: html_document
date: "MM-DD-YYYY"
---

#Source the data cleaning file to use correct dataset & source functions.

```{r setup, include=FALSE}
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)

knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Dataset: This is the dataset created from Step1.DataCleaning.R 
#Please ensure you use the correct data file, as naming conventions need to match.
valdata<-read.csv("FILEPATH/bcrpp_valdata.R")
```

```{r}
#Functions required
source("FILEPATH/Functions_ExistingModelValidation.R", local=knitr::knit_global())
```

```{r echo=FALSE, include=FALSE}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('dplyr'))
suppressWarnings(library("lubridate"))
suppressWarnings(library("gtools"))
suppressWarnings(library("ggplot2"))

#BCRAT packages - Load from R package available online
local({r <- getOption("repos")
       r["CRAN"] <- "https://cran.r-project.org" 
       options(repos=r)
})
```

# Step 1. Prepare dataset for iCARE model 

##  1a. Divide dataset by age
```{r}
#Add time of onset variables, divide dataset into <50, >=50

data_icare <- valdata %>%  
  mutate(time.of.onset=ifelse(case_5yinv==1 & !is.na(timediff_case_yr), timediff_case_yr, Inf),
         time.of.onset2=ifelse(case_5y==1 & !is.na(timediff_case_yr), timediff_case_yr, Inf),
         heightnew=height/10,
         age=round(age),
         agestart=round(age),
         ageend=round(age)+5)

#fix bbd and height for icare model
drops<-c("height")
data_icare<-data_icare[,(! names(data_icare) %in% drops),]
data_icare$height<-data_icare$heightnew
data_icare$pmh_type_icare_cont<-as.numeric(data_icare$pmh_type_icare)
data_icare<-rename(data_icare, bbd=BBD_history)
data_icare<-rename(data_icare, oc_current=ocuse_current)
```
```{r}
#Divide dataset
data_icare_lt50<-data_icare[which (data_icare$age<50),]
data_icare_ge50<-data_icare[which (data_icare$age>=50),]
```

##  1b. Fit names to iCARE objects
```{r}
#LT50: age_at_menarche, parity, age_first_birth, oc_ever, alcohol_intake, famhist, bmi_curc, oc_current, alcohol_intake, famhist, bmi_curc, bbd_hx, height

data_icare_lt50<- data_icare_lt50 %>% 
  mutate(age_at_menarche=agemenarche_cat7,
         parity=parity_cat,
         age_first_birth=ageflb_parous,
         oc_ever=ocuse_ever,
         alcohol_intake=alcgm_cat7,
         famhist=famhx_first,
         bmi_curc=ifelse(bmi_cat4==1, 0, 
                         ifelse(bmi_cat4==2, 1, 
                                ifelse(bmi_cat4==3, 2, 
                                       ifelse(bmi_cat4==4,3, NA)))))
```
#GE50: age_at_menarche, parity, age_first_birth, oc_ever, alcohol_intake, age_at_menopause, hrt, hrt_type, famhist, bmi_curc, alcohol_intake, bbd_hx, height
```{r}
data_icare_ge50<- data_icare_ge50 %>% 
  mutate(age_at_menarche=agemenarche_cat7,
         parity=parity_cat,
         age_first_birth=ageflb_parous,
         oc_ever=ocuse_ever,
         alcohol_intake=alcgm_cat7,
         age_at_menopause=meno_age_cat5_icare,
         famhist=famhx_first,
         hrt=hrt_use_icare,
         hrt_type=pmh_type_icare_cont,
         bmi_curc=bmi_cat3)

```

##  1c. Create iCARE datasets with needed variables only

```{r}
#Note: there cannot be any NA time of onset 
data_icare_lt50_val<-data_icare_lt50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","weight")]
#change names
names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.weights")

data_icare_lt50_val$sampling.weights<-1

data_icare_ge50_val<-data_icare_ge50[,c("subject_id", "cohort", "race","case_5yinv","age","ageend","time.of.onset","weight")]
#change names
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome","study.entry.age","study.exit.age","time.of.onset","sampling.weights")

data_icare_ge50_val$sampling.weights<-1
```


# Step 2. Calculate absolute risk using iCARE-Lit model

## 2a. installation
```{r echo=FALSE, include=FALSE}
install.packages("BiocManager")
BiocManager::install("iCARE")
library(iCARE)
```

```{r include=FALSE}
# ** Data notes: iCARE dependencies include: *model.lit.lt50*, *cov_info_lit_lt50*, *model.lit.ge50*, and *cov_info_lit_ge50* apply to ALL studies**
# ** For US based studies, you can use *bc_inc_US* and *mort_US* and *ref_lit_lt50_US* and *ref_lit_ge50_US***
# ** If you are a non-US cohort, PLEASE USE reference datasets, incidence rates, and mortality rates that are relevant to your country/region**
# ** You will need to replace reference to datasets in 2b and 2c to align with your data **
# ** Variable names included in ref. dataset need to be same names as those in the validation dataset. Change names in validation dataset if needed.**
```

##  2b. <50 absolute risk
```{r}
#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort

results_iCARE_lt50=computeAbsoluteRisk(model.formula = model.lit.lt50, model.cov.info = cov_info_lit_lt50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_lt50,
                                       model.ref.dataset = ref_lit_lt50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_lt50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_lt50[, names(ref_lit_lt50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)

```

## 2c. >=50 absolute risk
```{r}
#See above note: will need to replace reference datasets, mortality, and inc rates based on cohort

results_iCARE_ge50=computeAbsoluteRisk(model.formula = model.lit.ge50, model.cov.info = cov_info_lit_ge50, 
                                       model.snp.info = NULL, model.log.RR = beta_lit_ge50,
                                       model.ref.dataset = ref_lit_ge50_US, 
                                       model.ref.dataset.weights = NULL,
                                       model.disease.incidence.rates=bc_inc_US,
                                       model.competing.incidence.rates = mort_US,
                                       model.bin.fh.name = "famhist", n.imp = 5,
                                       apply.age.start=data_icare_ge50$age, apply.age.interval.length=5,
                                       apply.cov.profile  = data_icare_ge50[, names(ref_lit_ge50_US)], #dataset to apply 
                                       apply.snp.profile = NULL,
                                       use.c.code = 1,  return.lp = TRUE, 
                                       return.refs.risk = TRUE)
```

##2d. Store risk estimates in dataframes for calibration
```{r}
#store summary with abs risk as data frame so add in risk estimates
results_iCARE_lt50det<-as.data.frame(results_iCARE_lt50$details)

#add in lps
lps_lt50<-as.vector(results_iCARE_lt50$lps)
results_iCARE_lt50det$lps<-lps_lt50 

results_iCARE_ge50det<-as.data.frame(results_iCARE_ge50$details)
lps_ge50<-as.vector(results_iCARE_ge50$lps)
results_iCARE_ge50det$lps<-lps_ge50

#Dataset: results_iCARE_lt50det has Risk_Estimate- abs risk & lps 
#colnames(data_icare_lt50_val)
#Set absolute risk vector & lps
data_icare_lt50_val$absrisk.icare.lt50<-results_iCARE_lt50det$Risk_Estimate
data_icare_lt50_val$lps.icare.lt50<-results_iCARE_lt50det$lps

data_icare_ge50_val$absrisk.icare.ge50<-results_iCARE_ge50$risk
data_icare_ge50_val$lps.icare.ge50<-results_iCARE_ge50$lps
names(data_icare_lt50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights", "absrisk", "lp")
names(data_icare_ge50_val)<-c("id","cohort","race","observed.outcome", "study.entry.age", "study.exit.age", "time.of.onset", "sampling.weights", "absrisk", "lp")
#head(data_icare_lt50_val)
```

## 2e. Summarize estimates by age

```{r}
a<-summarizerisks(results_iCARE_lt50det, Risk_Estimate)
b<-summarizerisks(results_iCARE_lt50det, lps)
c<-summarizerisks(results_iCARE_ge50det, Risk_Estimate)
d<-summarizerisks(results_iCARE_ge50det, lps)
icarerisks<-rbind(a,b,c,d)
rownames(icarerisks)<-c("LT50,AbsRisk","LT50,LP","GE50,AbsRisk","GE50,LP")
icarerisks
```


# Step 3. Combine datasets

```{r}
data_icare_full<-rbind(data_icare_lt50_val, data_icare_ge50_val)
data_icare_full$absrisk<-as.numeric(data_icare_full$absrisk)
a<-summarizerisks(data_icare_full, absrisk)
b<-summarizerisks(data_icare_full, lp)
icarerisks<-rbind(a,b)
rownames(icarerisks)<-c("AbsRisk","LP")
icarerisks
```

# Step 4. Calibraiton
## 4a. E/O invasive
```{r}
eo_invas<-as.data.frame(obsexpected_overall(data_icare_full))
print("E/O invasive only")
eo_invas
```

```{r}
eo_bycat_invas<-as.data.frame(obsexpected_bycat_abs(data_icare_full))
print("E/O by cat, invasive")
eo_bycat_invas
```

##4b. Calibration plots

```{r}
icare_absrisk_byabs<-as.data.frame(returnabsrisk_abs(data_icare_full))
print("Absolute risk by absolute risk decile")
icare_absrisk_byabs
```

## Calibration plot of absolute risk
```{r}
a1<-suppressWarnings(plotcalibration_abs(icare_absrisk_byabs))
plot(a1)
```

## Calibration plot of absolute risk #2
```{r}
a1b<-suppressWarnings(plotcalibration_abs2(icare_absrisk_byabs))
plot(a1b)
```

# AUC for Gail model
```{r}
AUC<-as.data.frame(AUCcalc(data_icare_full))
print("AUC overall")
AUC
```

```{r}
AUCageinc<-as.data.frame(AUCcalc_ageinc(data_icare_full))
print("AUC age incorporated")
AUCageinc
```





